---
title: "Introduction to Data Science"
output: github_document
---

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
library(dplyr)
library(lubridate)

nice_date<-function(x){
  ##Nice print: take a date object and print it as "Monday, September 10"
  nice_date<-paste0(wday(x,label=TRUE,abbr=FALSE),
                  ", ",
                month(x,label=TRUE,abbr=FALSE),
                " ",
                day(x)
                )
  nice_date
}

## Day for first day of the week start sequence  (eg Mon or Tues)
first_week_days_start<-as_date("2017-08-28")

## Day for first day of the week end sequence  (eg Mon or Tues)
first_week_days_end<-as_date("2017-12-04")

## Sequence of days for the first day of the week (eg Mon or Tues)
first_week_days<-seq(first_week_days_start,first_week_days_end,by=7)

##Defines a sequence of days for the second day of the week (eg Weds or Thurs)
## if it's a second day start, then this should be the first in full weeks, then just push the first class as -7. 

second_week_days_start<-as_date("2017-08-30")

second_week_days_end<-as_date("2017-12-06")

second_week_days<-seq(second_week_days_start,second_week_days_end,by=7)

## Defines a sequence of days for the homeworks

#Start Day
hw_days_start<-as_date("2017-08-27")

#End day
hw_days_end<-as_date("2017-12-03")  

## Sequence of hw days
hw_days<-seq(hw_days_start,hw_days_end,by=7)

```

## Vanderbilt University 
## Human and Organizational Development 
## Course Number HOD 3200 
## Fall 2017

William R. Doyle 

Office: 207D Payne

Office Hours: Mondays and Wednesdays, 1-3 or by appointment

Email: [w.doyle@vanderbilt.edu](mailto:w.doyle@vanderbilt.edu)

Twitter: [\@wdoyle42](https://twitter.com/wdoyle42)

Phone: (615) 322-2904

Introduction
=============

We have entered a time in which vast amounts of data are more widely
available than ever before. At the same time, a new set of tools has
been developed to analyze this data and provide decision makers with
information to help them accomplish their goals. Those who engage with
data and interpret it for organizational leaders have taken to calling
themselves data scientists, and their craft data science. Other terms
that have come into vogue are "Big Data," "Predictive Analytics" and
"Data Mining." These can seem to be mysterious domains. The point of
this class is to demystify much of this endeavor for individuals who
will be organizational leaders.

The class is structured around developing students' skills in three
areas: getting data, analyzing data to make predictions, and
presenting the results of analysis. For each area, the subtopics are
as follows:

Getting Data Topics
--------------------
* Tools of the Trade: R and Rstudio, git and Github
* Working with pre-processed data and flat files
* Getting data from the web: webscraping, using forms, using Application Programming Interfaces
* Using databases

Analyzing Data Topics
----------------------
* Descriptives and conditional means
* Regression
* Supervised learning: classification
* Unsupervised learning: K-means and nearest neighbors clustering
* Evaluating multiple models

Presenting Data Analysis Topics
---------------------------------
* Descriptives: histograms, density plots, bar plots, dot
  plots
* Scatterplots
* Lattice graphics and small multiples
* Maps
* Interactive graphics

Evaluation
===========
Students will be evaluated based in two areas: weekly assignments and
the final project. 

* Problem sets: 65% Each week I will assign a problem set for
  students to complete. These problem sets will be assigned on Monday,
  and will be due the next Sunday night at 11:59:59 pm. No late
  assignments will be accepted. Each assignment will be graded on a
  100 point scale. Your lowest grade will be dropped. 

* Final Project 35%: During the course of the semester you will
  work on a final assignment utilizing your skills as a data
  analyst. We will discuss this assignment and my expectations in
  detail during the course of the semester. There will be four progress reports
  due for the final project, each of which will be worth 12.5% of the final  grade for the project. No late progress reports will be accepted. The final product will account for the remaining 50%. No late final products will be accepted. 

Texts
======

*Required Texts*

We will have two texts for the course. The first is Hadley Wickham's book, 
[R for Data Science](http://r4ds.had.co.nz/). Wickham is generously making this book available for free. However, I strongly encourage you to buy this book from [O'Reilly](http://amzn.to/2aHLAQ1).

[Amazon](https://www.amazon.com/R-Data-Science-Hadley-Wickham/dp/1491910399/ref=as_li_ss_tl?ie=UTF8&qid=1469550189&sr=8-1&keywords=R+for+data+science&linkCode=sl1&tag=devtools-20&linkId=6fe0069f9605cf847ed96c191f4e84dd)

The other text is Nate Silver's *Signal and the Noise*. 

Silver,N. (2012). *The signal and the noise: Why so many predictions fail-but some don’t*. New York: Penguin. 

[Amazon](https://www.amazon.com/Signal-Noise-Many-Predictions-Fail-but/dp/0143125087) 

[Half.com](http://product.half.ebay.com/The-Signal-and-the-Noise-Why-So-Many-Predictions-Fail-But-Some-Dont-by-Nate-Silver-2015-Paperback/203429545&tg=info) 

[Your local bookseller](http://www.indiebound.org/indie-bookstore-finder)

*Reserve*

I've placed three books by [Edward Tufte](https://www.edwardtufte.com/tufte/books_vdqi) on reserve for you. These are masterpieces in the area of visualizing quantitative information. You should take a look at these for ideas and inspiration--I've noted
the sections that are most helpful in various parts of the syllabus. 

Tufte, E. R. (1990). Envisioning information. Cheshire, CT: Graphics Press.

Tufte, E. R. (1997) Visual explanations.  Cheshire, CT: Graphics press.

Tufte, E. R.(2001). The visual display of quantitative information (2nd Edition). Cheshire, CT: Graphics press.


Lecture Notes
===============

My lecture notes are available at:  [willdoyle.us/data_step](http://willdoyle.us/data_step). These include both code and notes for the week. They will also be available in your private github repository. 

Web Resources
===============

When appropriate for each week, web resources are linked directly from 
the syllabus. You will find a wealth of resources online, including other versions of this class offered as Massive Online Open Courses. I encourage you 
to take full advantage of the wealth of online materials that are 
available. [Stack Overflow](https://stackoverflow.com/questions/tagged/r) is your friend, but search carefully for your question. It is VERY likely that your question has already been asked.  

Software
==========
We will use only free, [open source](https://opensource.org/osd-annotated) software in this course. 

We will use  [R](https://cran.r-project.org/), an open-source data analytic
platform for all analysis. R appears to be the most widely used data analysis software in data science. We will utilize [Rstudio](https://www.rstudio.com/home/) as our integrated development environment (IDE) for R. 

We will also use [git](https://git-scm.com/), a distributed version control program, and  [Github](https://github.com/),  an online hosting platform. [Github Desktop](https://desktop.github.com/) will serve as our Graphical User Interface to git and GitHub:  RStudio is fully integrated with git and Github, making it an ideal IDE for these purposes. Class assignments will be distributed through GitHub and will be collected and graded through GitHub as well. 

Communication
================
My office is in 207D Payne, and my phone number is (615) 322-2904. Please always feel free to stop by during office hours (Mondays and Wednesdays 12:30-3, right after class) or to call. If my office hours don't work for you, please make an appointment. Student communications, including emails are my priority. However, due to the volume of email I receive, I may miss your message. To help with this problem, please place the phrase ``HOD 3200'' in your subject line. I will search for these messages every time I access my email. You can also use Brightspace's email function, which will automatically do this for you. If you have a general question that I can answer for the whole class, send me a message on twitter at [\@wdoyle42](https://twitter.com/wdoyle42), tagged \#hoddatasci, or you can send a direct message. 

The Tedious Stuff
==================
I don't take attendance. This class will be impossible if you don't show up. It's reasonable to contact me if for some reason you can't make it for a given class session. 

We must use laptops in this class, despite the [substantial](http://journals.sagepub.com/doi/abs/10.1177/0956797614524581) [body](http://www.sciencedirect.com/science/article/pii/S0360131512002254?via%3Dihub) of [evidence](http://www.sciencedirect.com/science/article/pii/S0272775716303454) that laptop use in classrooms hinders learning. To mitigate this problem, the following standards will ALWAYS apply in class. You may have RStudio open. You may also have a web browser open to a web page that is relevant to course content. You MUST turn off all notifications and messaging programs. If your web browser is open to Facebook, Instagram or other purely social sites, I will ask you to leave class for the day.  

Mobile phone use is never appropriate in class. I will ask you to leave if you are using your mobile phone at any time. Exceptions are to be arranged BEFORE class, not when I observe you using your mobile phone. 

Honor Code Statement
===============

All assignments for this class, including weekly assignments and the final project, are to be conducted under the obligations set out in Vanderbilt's Honor Code. Please click [here](http://www.vanderbilt.edu/student_handbook/the-honor-system/)to review the honor code.

There will be two quite different standards for completing the
assignments and the final project. 

*Assignments* You may collaborate with anyone and you may utilize
  any resource you wish to complete these assignments. 
  
*Final Project* All of the work on the final assignment must be
  your own. Anyone's work that you reference should be cited, as
  usual. All data that you do not personally collect must be cited, as with any other resource. 

If you have any questions at all about the honor code or how it will
be applied, ask me right away. 

Schedule
=========
```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-1
```
## `r nice_date(second_week_days[i]-7)` Topic for the Week: Getting Data 1-- Tools of the Trade

*Resources*

Wickham: [Introduction](http://r4ds.had.co.nz/introduction.html), [Explore: Introduction](http://r4ds.had.co.nz/explore-intro.html), [Workflow: basics](http://r4ds.had.co.nz/workflow-basics.html), [Workflow: projects](http://r4ds.had.co.nz/workflow-projects.html)

Silver, Chapters 1-4

R Intro and Resources

[Download R](https://cran.r-project.org/)

[R Basics](https://www.datacamp.com/home)

[Download Rstudio](https://www.rstudio.com/products/rstudio/download3/)
You want the "Desktop" version, free license

[Rstudio Intro and Resources](https://www.rstudio.com/online-learning/)

[Download git](https://git-scm.com/downloads)

[Download GitHub Desktop](https://desktop.github.com)

[Github Intro and Resources](https://guides.github.com/activities/hello-world/)

## `r nice_date(first_week_days[i])`  Getting Data 1: Tools of the Trade

Subtopics: "verbs" of data wrangling, file types, working with git and GitHub. 

*Lesson Notes*

Chapter 1, Introduction: `01-intro.Rmd`.

## `r nice_date(second_week_days[i])`  Getting Data 1: Tools of the Trade

*Standing meeting*

*Lab Practical* R Basics, "verbs" of data wrangling

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## `r nice_date(first_week_days[i])` Topic for the Week: Analyzing Data 1-- Conditional Means

*Resources*
  
Wickham: [Data transformation](http://r4ds.had.co.nz/transform.html)  
Silver, Chapters 5-9, 12-13

*Lecture Notes*

Chapter 2, Condtional Means: `02-conditional_means.Rmd`.

*Assignments*

Assignment `r i-1` Due Midnight, `r nice_date(hw_days[i-1])`

## `r nice_date(second_week_days[i])` Conditional Means, continued

*Standing Meetings*

*Lab Practical: Conditional Means*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```


## `r nice_date(first_week_days[i])` Topic for the Week: Presenting Data 1-- Descriptives

Subtopics: bar plot, density plot, dot plots, histograms 

*Resources*

Wickham: [Data visualization](http://r4ds.had.co.nz/data-visualisation.html)

[Data transformation](http://r4ds.had.co.nz/exploratory-data-analysis.html)

[Cookbook for R:Bar and Line Graphs](http://www.cookbook-r.com/Graphs/Bar_and_line_graphs_%28ggplot2%29/)

[Cookbook for R: Plotting Distributions](http://www.cookbook-r.com/Graphs/Plotting_distributions_%28ggplot2%29/)

*Lecture Notes*

Chapter 3, Plotting Distributions and Condtional Means: `03-plot_means.Rmd`.

*Assignments*

Assignment `r i-1` Due Midnight `r nice_date(hw_days[i-1])`

## `r nice_date(second_week_days[i])` Descriptive Graphics, Continued

*Standing Meeting*
Progress reports due for section 1. 

*Lab Practical: Presenting results in graphical format: barplots, density plots, dot plots, histograms*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## `r nice_date(first_week_days[i])` Topic for the Week: Getting Data 2--pre-processed data, flat files, basic concepts of "tidy data"

*Resources*

Wickham: [Data import](http://r4ds.had.co.nz/data-import.html), [Tidy data](http://r4ds.had.co.nz/tidy-data.html)

*Lecture Notes*

Chapter 4, Flat Data `04-flat_data.Rmd`

*Assignments*

Assignment `r i-1` Due Midnight `r nice_date(hw_days[i-1])`

## `r nice_date(second_week_days[i])`  Pre-processed data, continued

*Standing Meeting*
Progress reports due for section 2.

*Lab Practical: working with various data formats*
```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## `r nice_date(first_week_days[i])` Topic for the Week: Analyzing Data 2--linear regression

*Resources*

Wickham: [Model: Introduction](http://r4ds.had.co.nz/model-intro.html),  [Model Basics](http://r4ds.had.co.nz/model-basics.html), [Model Building](http://r4ds.had.co.nz/model-building.html)

*Lecture Notes*

Chapter 5, Linear Regression `05-regression.Rmd`

*Assignments*

Assignment `r i-1` Due Midnight `r nice_date(hw_days[i-1])`

## September `r nice_date(second_week_days[i])` Linear Regression, continued

*Standing Meetings*

*Lab Practical: linear regression*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## `r nice_date(first_week_days[i])` Topic for the Week: Presenting data 2--scatterplots

*Resources*

Wickham: [Data Visualization](http://r4ds.had.co.nz/data-visualisation.html), [Graphics for Communication](http://r4ds.had.co.nz/graphics-for-communication.html)

Tufte, Visual Display chapters 4 and 5. 

Tufte, Envisioning Information, chapter 2

*Lecture Notes*

Chapter 6, Scatterplots `06-scatterplots.Rmd`

*Assignments*

Assignment `r i-1` Due Midnight `r nice_date(hw_days[i-1])`

## `r nice_date(second_week_days[i])` Scatterplots, continued

*Standing Meetings*
Progress reports due for section 1. 

*Lab Practical: Presenting Data via Scatterplots*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## `r nice_date(first_week_days[i])` Topic for the week: Getting Data 3-- Getting data from the web

*Resources*

Rvest Vignette: [https://cran.r-project.org/web/packages/rvest/vignettes/selectorgadget.html](https://cran.r-project.org/web/packages/rvest/vignettes/selectorgadget.html)

[Reed College rvest introduction](http://www.reed.edu/data-at-reed/resources/R/rvest.html)

[rvest tutorial](https://stat4701.github.io/edav/2015/04/02/rvest_tutorial/)

*Lecture Notes*

Chapter 7, Web Scraping and APIs, `07-webscrape.Rmd`

*Assignments*

Assignment `r i-1` Due Midnight `r nice_date(hw_days[i-1])`

## `r nice_date(second_week_days[i])` Web data, continued

*Standing Meetings*
Progress reports due for section 2. 

*Lab Practical*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```
## `r nice_date(first_week_days[i])` Topic for the Week: Analyzing Data 3--Classification

*Resources*

James, G., Witten, D., Hastie, T., & Tibshirani, R. (2013). An introduction to statistical learning (Vol. 6). New York: Springer. [Chapter 4](http://www-bcf.usc.edu/~gareth/ISL/ISLR%20Sixth%20Printing.pdf) , 
[Chapter 4 Lab R Code](http://www-bcf.usc.edu/~gareth/ISL/Chapter%204%20Lab.txt)

Althoff, T., Danescu-Niculescu-Mizil, C., & Jurafsky, D. (2014, May). How to Ask for a Favor: A Case Study on the Success of Altruistic Requests. In ICWSM. (Available Here)[http://www.aaai.org/ocs/index.php/ICWSM/ICWSM14/paper/download/8106/8101]

*Lecture Notes*

Chapter 8, Classification, `08-classification.Rmd`

*Assignments*

Assignment `r i-1` Due Midnight `r nice_date(hw_days[i-1])`

##`r nice_date(second_week_days[i])` Classification, continued

*Standing Meetings*

*Lab Practical*

Classifying behavior via text analysis: random acts of pizza.

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## `r nice_date(first_week_days[i])` Topic for the Week: Presenting Data 3-- Lattice plots and small multiples

*Resources*

Tufte, Envisioning Information, Chapter 4

*Lecture Notes*

Chapter 9, Small Multiples `09-small_multiples.Rmd`

*Assignments*

Assignment `r i-1` Due Midnight `r nice_date(hw_days[i-1])`

## `r nice_date(second_week_days[i])` Lattice Plots and Small Multiples, continued.

*Standing Meetings*

Progress reports due for section 1. 

*Lab Practical*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## `r nice_date(first_week_days[i])` Topic for the Week: Getting Data 4--databases and relational data

*Resources*
Wickham [Relational Data](http://r4ds.had.co.nz/relational-data.html)

Working with Databases in R, available: [https://cran.r-project.org/web/packages/dplyr/vignettes/databases.html](https://cran.r-project.org/web/packages/dplyr/vignettes/databases.html)

*Lecture Notes*

Chapter 10, Databases `databases.Rmd`

*Assignments*

Assignment `r i-1` Due Midnight `r nice_date(hw_days[i-1])`

## `r nice_date(second_week_days[i])` Databases, continued

*Standing Meetings*

*Lab Practical* Databases and relational data, collaborating via github

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```
## `r nice_date(first_week_days[i])` Topic for the Week: Training and testing multiple models

*Resources*

Wickham [Many Models](http://r4ds.had.co.nz/many-models.html)

*Lecture Notes*

Chapter 13, Multiple Models `multiple_models.Rmd`

*Assignments*

Assignment `r i-1` Due Midnight `r nice_date(hw_days[i-1])`

## `r nice_date(second_week_days[i])` Training and testing, continued

*Standing Meeting*

*Lab Practical*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## `r nice_date(first_week_days[i])` Topic for the Week: Presenting Data 4--Maps

*Resources*

Lovelace, Cheshire et al (2015) Introduction to visualising spatial data in R. [Online](https://cran.r-project.org/doc/contrib/intro-spatial-rl.pdf)

*Lecture Notes*

Chapter 12, Mapping: `12-mapping.Rmd`

*Assignments*

Assignment `r i-1` Due Midnight `r nice_date(hw_days[i-1])`

## `r nice_date(second_week_days[i])` Mapping, continued

*Standing Meetings*

Progress reports due for section 1 and 2. 

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## `r nice_date(first_week_days[i])`  `r nice_date(second_week_days[i])`

Thanksgiving Break: No Class, but lots of data analysis to do!

*Lecture Notes*

Take a look at Chapter 14, Interactive Graphics: `14-interactive_graphics.Rmd` This will be very helpful for creating nifty graphics for your final assignment. 

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## `r nice_date(first_week_days[i])` Topic for the Week: Analyzing Data 4--Unsupervised learning 

Subtopics: k-means, nearest neighbor clustering

*Resources*

James, G., Witten, D., Hastie, T., & Tibshirani, R. (2013). An introduction to statistical learning (Vol. 6). New York: Springer. [Chapter 10](http://www-bcf.usc.edu/~gareth/ISL/ISLR%20Sixth%20Printing.pdf) , 
[Chapter 10 Lab R Code](http://www-bcf.usc.edu/~gareth/ISL/Chapter%2010%20Labs.txt)

*Lecture Notes*

Chapter 11, Unsupervised Learning `11-unsupervised.Rmd`

*Assignments*

Assignment `r i-2` Due Midnight `r nice_date(hw_days[i])`

## `r nice_date(second_week_days[i])` Unsupervised Learning, continued

*Standing Meetings*

*Lab Practical:* K-means clustering, nearest neighbor classification


```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## `r nice_date(first_week_days[i])` Class Presentations

Section 1

*Assignments*

Assignment `r i-2` Due Midnight `r nice_date(hw_days[i])`

## `r nice_date(second_week_days[i])` Class Presentations 

Section 2

## Final Projects Due `r nice_date(first_week_days[i]+7)`, midnight