---
output: github_document
---

# Welcome to Data Science!

Today, we'll be working on getting you set up with the tools you will need for this class. Once you are set up, we'll do what we're here to do: analyze data!

Here's what we need to get done today:

1.  Introductions  
2.  Installing R
3.  Installing RStudio
4.  Installing git
5.  Getting set up on GitHub
6.  Initializing everyone's GitHub repos
7.  Starting assignment 1, "hello, world!"

## Introductions

We need three basic sets of tools for this class. We will need `R` to analyze data. We will need `RStudio` to help us interface with R and to produce documentation of our results. Last, we will need `git` and GitHub to communicate our results to the wider world.

## Installing R

R is going to be the only programming language we will use. R is an extensible statistical programming environment that can handle all of the main tasks that we'll need to cover this semester: getting data, analyzing data and communicating data analysis.

If you haven't already, you need to download R here: <https://cran.r-project.org/>.

## Installing RStudio

When we work with R, we communicate via the command line. To help automate this process, we can write scripts, which contain all of the commands to be executed. These scripts generate various kinds of output, like numbers on the screen, graphics or reports in common formats (pdf, word). Most programming languages have several **I** ntegrated **D** evelopment **E** nvironments (IDEs) that encompass all of these elements (scripts, command line interface, output). The primary IDE for R is RStudio.

If you haven't already, you need to download RStudio here: <https://rstudio.com/products/rstudio/download/>. You need the free RStudio desktop version.

## Installing git

`git` is a version control program. A standard problem in programming is how to track the changes that have been made to a project. These challenges are well-known to anyone who has had to work on a complex document, particularly with collaborators. Because these problems are particularly severe in programming, they developed a set of tools called version control. These will keep track of every change you make to a file, and record why you made the change. Download git here and accept all of the defaults on installation: (<https://git-scm.com/downloads>)

## Getting set up on GitHub

GitHub is online hosting service that is widely used by programmers. It allows you to easily share your work with the world. It is integrated with git, so version control is easy to do. If you haven't already, go to (<https://GitHub.com>) and sign up for free to get a username. Share the username with me as soon as you have one.

## Initializing repos

Everyone in the class will need a repository (repo from now on) on our GitHub organization. All I need from you is your username on GitHub, then I can add your repo to our organization.

## Installing GitHub Desktop

GitHub desktop is a **G** raphical **U** ser **I** nterface for git. We won't use it much, but you will need it to "clone" your git repository.

Download GitHub desktop (<https://desktop.GitHub.com/>). Choose the beta version, which appears to be working nicely. Accept all of the defaults on installation.

Unzip and open up the GitHub app. Then using the app, sign into GitHub using your username and ID.

## Integrating RStudio with GitHub

To integrate RStudio with GitHub, you need to enable git as your version control within RStudio. Go to Preferences--\>Git/SVN and make sure the "enable version control" box is checked. Make sure that you can see a path to git.

Further instructions (if needed) are [here](https://support.rstudio.com/hc/en-us/articles/200532077-Version-Control-with-Git-and-SVN).

## Cloning down your repo

First in RStudio, you'll need to generate an SSH key. This will let GitHub know it's safe to talk to your computer. In RStudio, go to Preferences--\> Git/SVN, then click on "Create RSA Key". Once that's done, you can click on the blue text that says "View public key." Copy all of the text in the public key.

Now go to [GitHub](https://github.com) and in the upper right hand corner by your avatar, click on Settings. Go to SSH and GPG keys. Click "New SSH Key" and paste in the text you copied from RStudio. Name the key, and you should be all set.

Now, go to your repo for this class. Everyone's repo is in the `hoddatasci` organization, and uses the naming convention `student_<yourlastname>`. Click the green button that says "clone or download," then copy the link provided by clicking on the clipboard next to the link.

*N.B. in this class, and generally in programming, when you see* `<text>` *that means that you need to substitute something in.*

Now go to GitHub desktop, and click the "clone a repository." Paste in the link you just copied, choose a location on your computer, and proceed. Choose a good spot on your computer for the project files-- wherever you usually keep class directories. DO NOT PUT YOUR DIRECTORY ON YOUR DESKTOP OR IN DOWNLOADS. This project will be the only place you need to work for this class the entire semester.

## Linking GitHub and RStudio

From RStudio, you'll need to click New Project--Existing Directory-- then choose the directory where you just downloaded the GitHub repository.

Name the project "central".

Once you click "create project," you should see a "git" tab in your environment. Open up the file named `01-intro.Rmd` and take a look.

## Yes We Code! Running R Code

The following code chunks will be our first use of R in this class. We're going to grab some data that's part of the [college scorecard](https://collegescorecard.ed.gov/data/documentation/) and do a bit of analysis on it.

## .Rmd files

Open the `01-intro.Rmd` file. In RStudio, go to File--\>Open, then find the `01-intro.Rmd` file in the directory.

.Rmd files will be the only file format we work in this class. .Rmd files contain three basic elements:

1.  Script that can be interpreted by R.
2.  Output generated by R, including tables and figures.  
3.  Text that can be read by humans.

From a .Rmd file you can generate html documents, pdf documents, word documents, slides . . . lots of stuff. All class notes will be in .Rmd. All assignments will be turned in as .Rmd files, and your final project? You guessed it, .Rmd.

In the `01-Intro.Rmd` file you'll notice that there are three open single quotes in a row, like so: ```` ``` ```` This indicates the start of a "code chunk" in our file. The first code chunk that we load will include a set of programs that we will need all semester long.

## Using R Libraries

When we say that R is extensible, we mean that people in the community can write programs that everyone else can use. These are called "packages." In these first few lines of code, I load a set of packages using the library command in R. The set of packages, called `tidyverse` were written by Hadley Wickham and others and play a key role in his book. To install this set of packages, simply type in `install.packages("tidyverse")` at the R command prompt.

To run the code below in R, you can:

-   Press the "play" button next to the code chunk
-   In OS X, place the cursor in the code chunk and hit `CMD+RETURN`
-   In Windows, place the cursor in the code chunk and hit `CTRIL+RETURN`

```{r}
## Get necessary libraries-- won't work the first time, because you need to install them!
library(tidyverse)
```

## Loading Datasets

Now we're ready to load in data. The data frame will be our basic way of interacting with everything in this class. The `sc_debt.Rds` data frame contains information from the college scorecard on different colleges and universities.

However, we first need to make sure that R is looking in the right place. When you opened up your project, RStudio automagically took you to the directory for that project. But because we keep lessons in a separate directory, we need to point R to the right place. This is called setting the working directory, and can be done either by using the command `setwd` or in RStudio by going to "Session--\>Set Working Directory--\>Choose Directory." Choose the directory where the file currently resides on your computer. Make sure to always set the working directory at the beginning of each session---not doing so causes a lot of headaches for new users.

```{r}
df<-readRDS("sc_debt.Rds") 
names(df)
```

| Name           | Definition                                                                                                                                                                                             |
|----------------|--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| unitid         | Unit ID                                                                                                                                                                                                |
| instnm         | Institution Name                                                                                                                                                                                       |
| stabbr         | State Abbreviation                                                                                                                                                                                     |
| grad_debt_mdn  | Median Debt of Graduates                                                                                                                                                                               |
| control        | Control Public or Private                                                                                                                                                                              |
| region         | Census Region                                                                                                                                                                                          |
| preddeg        | Predominant Degree Offered: Associates or Bachelors                                                                                                                                                    |
| openadmp       | Open Admissions Policy: 1= Yes, 2=No,3=No 1st time students                                                                                                                                            |
| adm_rate       | Admissions Rate: proportion of applications accepted                                                                                                                                                   |
| ccbasic        | Type of institution-- see [here](https://data.ed.gov/dataset/9dc70e6b-8426-4d71-b9d5-70ce6094a3f4/resource/658b5b83-ac9f-4e41-913e-9ba9411d7967/download/collegescorecarddatadictionary_01192021.xlsx) |
| selective      | Institution admits fewer than 10 % of applicants, 1=Yes, 0=No                                                                                                                                          |
| research_u     | Institiution is a research university 1=Yes, 0=No                                                                                                                                                      |
| sat_avg        | Average Sat Scores                                                                                                                                                                                     |
| md_earn_wne_p6 | Average Earnings of Recent Graduates                                                                                                                                                                   |
|ugds  | Number of undergraduates |

## Looking at datasets

We can look at the first few rows and columns of `df` by typing in the data name.

```{r look_at_data}
df
```

We can look at the whole dataset using View. Just delete the `#` sign below to make the code work. That `#` sign is a comment in R code, which indicates to the computer that everything on that line should be ignored.

```{r, eval=FALSE}
#View(df)
```

You'll notice that this data is arranged in a rectangular format, with each row showing a different college, and each column representing a different characteristic of that college. Datasets are always structured this way--- cases (or units) will form the rows, and the characteristics of those cases-- or variables--- will form the columns. Unlike working with spreadsheets, this structure is always assumed for datasets.

## Filter, Select, Arrange

In exploring data, many times we want to look at smaller parts of the dataset. There are three commands we'll use today that help with this.

\-`filter` selects only those cases or rows that meet some logical criteria.

\-`select` selects only those variables or columns that meet some criteria

\-`arrange` arranges the rows of a dataset in the way we want.

For more on these, please see this [vignette](https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html).

Let's grab just the data for Vanderbilt, then look only at the average test scores and admit rate. We can use filter to look at all of the variables for Vanderbilt:

```{r filter_select_arrange}
df%>%
  filter(instnm=="Vanderbilt University")
```

What's that weird looking `%>%` thing? That's called a pipe. This is how we chain commands together in R. Think of it as saying "and then" to R. In the above case, we said, take the data *and then* filter it to be just the data where the institution name is Vanderbilt University.

Many times, though we don't want to see everything, we just want to choose a few variables. `select` allows us to select only the variables we want. In this case, the institution name, its admit rate, and the average SAT scores of entering students.

```{r}
df%>%
  filter(instnm=="Vanderbilt University")%>%
  select(instnm,adm_rate,sat_avg)
```

`filter` takes logical tests as its argument. The code `insntnm=="Vanderbilt University"` is a logical statement that will be true of just one case in the dataset-- when institution name is Vanderbilt University. The `==` is a logical test, asking if this is equal to that. Other common logical and relational operators for R include

-   `>`, `<`: greater than, less than
-   `>=`, `<=`: greater than or equal to, less than or equal to
-   `!` :not, as in `!=` not equal to
-   `&` AND
-   `|` OR

Next, we can use `filter` to look at colleges with low admissions rates, say less than 10% ( or .1 in the proportion scale used in the dataset).

```{r}
df%>%
  filter(adm_rate<.1)%>%
  select(instnm,adm_rate,sat_avg)%>%
  arrange(sat_avg,adm_rate)%>%print(n=20)
```

Now let's look at colleges with low admit rates, and order them by SAT scores (`-sat_avg` gives descending order).

```{r}
df%>%
  filter(adm_rate<.1)%>%
  select(instnm,adm_rate,sat_avg)%>%
  arrange(-sat_avg)
```

And one last operation: all colleges that admit between 20 and 30 percent of students, looking at their SAT scores, earnings of attendees six years letter, and what state they are in, then arranging by state, and then SAT score.

```{r}
df%>%
  filter(adm_rate>.2&adm_rate<.3)%>%
  select(instnm,sat_avg,grad_debt_mdn,stabbr)%>%
  arrange(stabbr,-sat_avg)%>%
  print(n=20)
```

*Quick Exercise* Choose a different college and two different things about that college. Have R print the output.

## Summarizing Data

To summarize data, we use the summarize command. Inside that command, we tell R two things: what to call the new object (a data frame, really) that we're creating, and what numerical summary we would like. The code below summarizes median debt for the colleges in the dataset by calculating the average of median debt for all institutions.

```{r summarize}
df%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE))

df%>%
  summarize(median_debt=median(grad_debt_mdn,na.rm=TRUE))


```

*Quick Exercise* Summarize the average entering SAT scores in this dataset.

## Combining Commands

We can also combine commands, so that summaries are done on only a part of the dataset. Below, we summarize median debt for selective schools, and not very selective schools.

```{r combining commands}
df%>%
  filter(adm_rate<.1)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE))
```

What about for not very selective schools?

```{r}
df%>%
  filter(adm_rate>.3)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE))
```

*Quick Exercise* Calculate average earnings for schools where SAT\>1200

## Grouping Data

Another powerful tool is being able to calculate characteristics for various groups. For example, what are the average earnings for the three different types of colleges (public, private non-profit, private for-profit) in the dataset?

```{r grouping}
df%>%
  group_by(control)%>%
  summarize(mean_earnings=mean(md_earn_wne_p6,na.rm=TRUE))
```

*Quick exercise* Calculate average admission rate by type of college.

## Plotting Data

The last basic tool for looking at a dataset is plotting the data. The code below creates a scatterplot of admission rates by average SAT scores

```{r plotting}
## Plotting: bivariate
gg<-ggplot(data=df,aes(x=sat_avg,y=adm_rate))
gg<-gg+geom_point()
gg<-gg+geom_smooth()
gg
```

*Quick exercise* Replicate the above plots, but debt level on the y axis.

## GitHub: save, stage, commit, push

When working with files in your directory, there are several basic steps you will take when using Git as your version control.

-   You should pull at the beginning of every session. You can do this by clicking pull--the down arrow-- in the upper right hand corner of the git tab in RStudio.

-   Saving the file means it is only available to you, on your computer. You should save files as you are working on them.

-   Staging the file means that you would like git (and GitHub) to keep track of the changes you're making to the file. Stage the file by clicking the "staged" box next to it in the Git tab in RStudio.

-   Committing means that you would like to keep the version of the file you generated. Think of this like emailing it to group members. It doesn't mean that it's done, but it does mean it's in a state that you would like to have a record of as you move forward. To commit a file in RStudio, click the "commit" button. You will be prompted to add a commit message. There's actually quite a lot of thought about what goes into a commit message. The best idea for now is simply to state why you did what you did, and avoid profanity or any demeaning language.

-   Pushing means that you will send the file and the record of all of the changes made to the file to GitHub. To push, click the "push" button in the Git tab in RStudio. You should do this every time you finish a working session, at an absolute minimum.

So every session should go like this:

pull--\>do work--\>save--\>stage(if new file)--\>commit

By far the most useful guide to working with R and git/GitHub is [Jenny Bryan's guide](https://happygitwithr.com/).

## Your first commit: Hello, World!

For today, I want you to create a file called `01-assignment_<lastname>.Rmd` in your GitHub repo for assignments. It should contain the following elements:

1.  A sentence that says "Hello, World"
2.  R output that summarizes one of the variables in the `sc_debt.Rds` dataset
3.  R output that shows a scatterplot for two of the variables in the `sc_debt.Rds` dataset.

Lucky for you this is is also your first assignment! Submit it under assignments, using the format `01-assignment_<lastname>.Rmd`. All assignments should be turned in using this format. Since my last name is Doyle, I would use `01-assignment_doyle.Rmd` as my file name. Unless your name is also Doyle, you should use a different name.

*Stretch Items*

If you have extra time, you can do the following:

1.  Calculate the average earnings for individuals at the most selective colleges, then compare that with individuals at the least selective colleges in the dataset.

2.  Find a way to determine whether colleges with very high SAT scores tend to have higher or lower debt than colleges with low SAT scores.

3.  Plot the relationship between SAT and debt. What do you see? Does this surprise you?

4.  Now, provide separate plots for SAT and debt by control of the institution.

Save, commit and push the assignment file to the assignments directory.
