---
title: "Conditional Means"
output:
  word_document: default
  html_document: default
always_allow_html: yes
---

# Analyzing Data: Conditional Means

The conditional mean will be our first predictive algorithm. Conditional means answer the question: "Given what we know about a certain case, what can expect to see, on average?" The conditional mean is a powerful tool that is typically quite easy to explain to decision-makers.

We'll go through the following steps:

1.  Computing and unconditional means
2.  Computing conditional means using a single predictor and calculating our error.
3.  Computing conditional means using multiple predictors and calculating our error.

## Motivating Example

Suppose you're talking to a high school senior about college choice and student debt. They want to know which types of colleges have graduates with higher and lower levels of debt. They're concerned about the amount of debt they will incur on the path to graduation. Essentially they're asking for a prediction: based on the characteristics of the college they attend, what will be the likely amount of debt?

We'll use data from the [college scorecard](https://collegescorecard.ed.gov/) to answer these questions.

## Libraries

We'll use `tidyverse` as usual, and we'll need the `yardstick` library for some calculations.

```{r}
library(tidyverse)
library(yardstick)
```

## Dataset for this week

This week we'll use another dataset from the college scorecard, this one focused on student debt. Here are the variable names and definitions:

```{r}
df<-readRDS("sc_debt.Rds") 
names(df)
```

| Name           | Definition                                                                                                                                                                                                |
|----------------|-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| unitid         | Unit ID                                                                                                                                                                                                   |
| instnm         | Institution Name                                                                                                                                                                                          |
| stabbr         | State Abbreviation                                                                                                                                                                                        |
| grad_debt_mdn  | Median Debt of Graduates                                                                                                                                                                                  |
| control        | Control Public or Private                                                                                                                                                                                 |
| region         | Census Region                                                                                                                                                                                             |
| preddeg        | Predominant Degree Offered: Associates or Bachelors                                                                                                                                                       |
| openadmp       | Open Admissions Policy: 1= Yes, 2=NO,3=No 1st time students                                                                                                                                               |
| adm_rate       | Proportion of applicants admitted                                                                                                                                                                         |
| ccbasic        | Carnegie classification, see [here](https://data.ed.gov/dataset/9dc70e6b-8426-4d71-b9d5-70ce6094a3f4/resource/658b5b83-ac9f-4e41-913e-9ba9411d7967/download/collegescorecarddatadictionary_01192021.xlsx) |
| sat_avg        | Average SAT scores                                                                                                                                                                                        |
| md_earn_wne_p6 | Median earnings of recent graduates                                                                                                                                                                       |
| ugds           | Number of undergraduates                                                                                                                                                                                  |
| selective      | Institution admits fewer than 10 % of applicants, 1=Yes, 0=No                                                                                                                                             |
| research_u     | Institiution is a research university 1=Yes, 0=No                                                                                                                                                         |

## Dependent Variable

The dependent variable for us will be *institutional-level* median debt for graduates, as reported by the college scorecard.

## Unconditional Means

If you were asked to predict the level of debt for this student without any information about the college itself, the best guess would be the overall average, or what we're going to call the unconditional mean.

```{r}
df%>%
  summarize(mean_debt_uncond=mean(grad_debt_mdn,na.rm=TRUE))
```

This gives us the overall average, and would be our prediction about the level of college debt the student could expect, if we didn't have any other information about the college.

## Unconditional Mean as a Predictor

Using the mean of a variable as a predictor is something we do all the time. In this example, our best guess for the level of debt incurred by a graduate is just the average level of debt across all institutions. We're going to add this guess to the dataset as our first prediction, using the `mutate` command.

```{r}
df<-df%>%
  mutate(mean_debt_uncond=mean(grad_debt_mdn,na.rm=TRUE))
```

Notice how this code is the same as above, except for the overwrite `df<-df%>%` and the use of `mutate` as opposed to `summarize.`

This is of course a terrible prediction. In the absence of any other information, it's many times the best we can do, but we really ought to be able to do better.

## Root Mean Squared Error

To understand how far off we are, we need to summarize our errors. We will use different ways of doing this this semester, but let's start with a very standard one, Root Mean Squared Error, or RMSE. An error term is the distance between each point and its prediction. We square the errors, take the average, then take the square root of the result. The name RMSE is exactly what RMSE is-- neat, huh?

$$ RMSE(\\hat{Y})=\\sqrt{ 1/n \\sum\_{i=1}^n(Y_i-\\hat{Y_i})^2} $$

Luckily we have an r function that can do this for us. The `rmse` function from the `yardstick` library will calculate the rmse for a dataset. We need to give it the dataset and the two columns we're interested, in this case actual debt `grad_debt_mdn` and our prediction `mean_debt_uncond`.

```{r}
rmse_uncond<-df%>%rmse(grad_debt_mdn,mean_debt_uncond)
rmse_uncond
```

This tells us that if we use the unconditional mean, we'll be off by an average of `r round(rmse_uncond$.estimate)`. Is that good or bad? We don't know! RMSE does not have a scale that tells us what it means. We have to understand it in the context of the problem we're working on. Lower is always better when it comes to RMSE, as a lower RMSE means a more accurate prediction.

## Conditional Mean With One Predictor

Public and private colleges in the US charge very different prices, and so it's plausible that graduates of these two types of institutions would have different levels of debt. We can calculate the conditional mean by first grouping by control using `group_by(control)` and then calculating the mean for each of the two groups: public or private.

```{r}
df%>%
  group_by(control)%>%
  summarize(mean_debt_control=mean(grad_debt_mdn,na.rm=TRUE))%>%
  arrange(control)
```

We can see that graduates of private institutions have higher debt levels than graduates of public institutions. We'll follow the same steps as we did with the unconditional mean and add the conditional mean to the dataset as a predictor.

```{r}
df<-df%>%
  group_by(control)%>%
  mutate(mean_debt_control=mean(grad_debt_mdn,na.rm=TRUE))%>%
  ungroup()
```

I added one more step: I ungrouped the dataset using `ungroup()`. It's good practice to do this at the end of an code chunk that uses the group function, otherwise other steps might be grouped in ways that were unintended.

We can then calculate the RMSE using our new prediction based on the conditional mean:

```{r}
rmse_control<-df%>%rmse(grad_debt_mdn,mean_debt_control)
rmse_control
```

Using the predictor increased our predictive accuracy considerably, from `r round(rmse_uncond$.estimate)` to `r round(rmse_control$.estimate)`!

## Conditional Mean With Two Predictors

It's also probably the case that the degree level for the institution matters. It takes longer to get a bachelor's degree than an associate degree, so we could expect that institutions that mostly give bachelor's degrees would have higher debt levels. To figure this out we need to group by control and predominant degree level.

Grouping with multiple predictors works exactly the same way as with one predictor. We just add the additional variable to the `group_by` command, like so:

```{r}
df%>%
  group_by(control,preddeg)%>%
  summarize(mean_debt_control_degree=mean(grad_debt_mdn,na.rm=TRUE))%>%
  arrange(mean_debt_control_degree)
```

Notice that I used `arrange` which will sort the dataset in descending order based on the variable selected, in this case average debt levels by control and degree awarded. We can see the very different levels of debt incurred at these four different types of institutions. Now let's add this conditional mean to our dataset as a prediction.

```{r}
df<-df%>%
  group_by(control,preddeg)%>%
  mutate(mean_debt_control_degree=mean(grad_debt_mdn,na.rm=TRUE))%>%
  ungroup()
```

```{r}
rmse_control_degree<-df%>%rmse(grad_debt_mdn,mean_debt_control_degree)
rmse_control_degree
```


Adding this variable has again improved our estimate.

## Conditional Mean With Three Predictors

We can continue this logic indefinitely, but let's focus on one more variable: the region of the country. Unlike the previous two variables, this one has eight different levels. Adding it as another grouping variable means we're going to end up with 2x2x8=32 different conditional means, one for each type of institution (public or private, 2 year or 4 year) in each region of the country

```{r}
df%>%
  group_by(control,preddeg,region)%>%
  summarize(mean_debt_control_degree_region=mean(grad_debt_mdn,na.rm=TRUE))%>%
  arrange(mean_debt_control_degree_region)%>%
  print(n=100)
```

Because it's so many more conditional means, I added the `print(n=100)` to the code chunk, so we can see all of the results.

Let's add these as predictors to the dataset:

```{r}
df<-df%>%
  group_by(control,preddeg,region)%>%
  mutate(mean_debt_control_degree_region=mean(grad_debt_mdn,na.rm=TRUE))%>%
  ungroup()
```

And calculate the RMSE:

```{r}
rmse_control_degree_region<-df%>%rmse(grad_debt_mdn,mean_debt_control_degree_region)
rmse_control_degree_region
```

This is a much better RMSE than what we started with. Is it good enough? Who knows? Again this depends on the context. But we can imagine the high school senior would have much more accurate information with which to make their decision.

```{r}
df%>%
  group_by(stabbr)%>%
  summarize(mean_debt_state=mean(grad_debt_mdn,na.rm=TRUE))%>%
  arrange(mean_debt_state)%>%
  ggplot(aes(x=fct_reorder(as_factor(stabbr),mean_debt_state),
                           y=mean_debt_state,fill=stabbr))+
  geom_bar(stat="identity")+
  coord_flip()+
  theme(legend.position="none")

```

```{r}
df<-df%>%
  group_by(stabbr,control,preddeg)%>%
  mutate(mean_debt_state=mean(grad_debt_mdn,na.rm=TRUE))%>%
  ungroup()
```

```{r}
rmse_state<-df%>%rmse(grad_debt_mdn,mean_debt_state)
rmse_state
```


```{r}
df<-df%>%
  select(-mean_debt_state)
```



## Applications of the Conditional Mean

When might we use the conditional mean?

-   Calculating average sales for a retail location by day of the week and month
-   Calculating yield rate (proportion of admitted students who attend) by geographic region and income level for a college.
-   Calculating average employee turnover by level of education and gender
