---
output: github_document
---
# Presenting Data 1: Plotting Conditional Means

The idea when plotting conditional means is to show how the outcome, or variable of interest, varies as a function of predictors. 

Today we'll be working with the college scorecard dataset, returning to our idea of predicting debt levels as a function of the characteristics of the college. 

## Setup for plotting conditional means

We start with a standard set of setup commands. Today we'll be working with `tidyverse`,us usual. To handle colors, we'll need the package `RColorBrewer.`

```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RColorBrewer)
```

Next we load in the data. 

## Loading Data 
```{r}
df<-readRDS("sc_debt.Rds")
```

Today, our primary outcome of interest will be student debt. We can quickly summarize this variable using `summarize` and functions like `mean` and `sd`.

```{r}
df%>%summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE),
               sd_debt=sd(grad_debt_mdn,na.rm=TRUE))
```


##Univariate Graphics

Univariate graphics help us understand what individual variables look like-- how are they distributed across the sample? Here's a quick rundown on some univariate graphics. Say we wanted a quick count of public and private institutions. We can use geom_bar to get this done. 

```{r}
df%>%
  group_by(control)%>%
  count()%>%
  ggplot(aes(x=control,y=n,fill=as_factor(control)))+
  geom_bar(stat="identity")
``` 

The next univariate graphic you should know is for continuous variables. The first thing you generally want is a histogram. 

```{r}
df%>%
  ggplot(aes(x=grad_debt_mdn))+
  geom_histogram()

``` 
Density plots provide a continuous graphic of the distribution of a variable:
```{r}
df%>%
  ggplot(aes(x=grad_debt_mdn))+
  geom_density(fill="lightblue")
```


## Predicting Student Debt

Our first prediction will calculate student debt as a function of control. 

```{r}
df%>%
  group_by(control)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE))

```

We can then take those results and plot them: 

```{r}
df%>%
  group_by(control)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE))%>%
  ggplot(aes(x=control,y=mean_debt))+
  geom_bar(stat="identity",position="dodge")
```

To make the bars in different colors we use the "fill" aesthetic:

```{r}
df%>%
  group_by(control)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE))%>%
  ggplot(aes(x=control,y=mean_debt,fill=control))+
  geom_bar(stat="identity",position="dodge")

```



*Quick Exercise: Create a bar plot showing average debt level by degree (2yr/4yr) instead of control*


```{r}
df%>%
  group_by(preddeg)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE))%>%
  ggplot(aes(x=preddeg,y=mean_debt,fill=preddeg))+
  geom_bar(stat="identity",position="dodge")
```




## Dot Plots

A dot plot can be a good way of displaying conditional means as well. Many times dot plots are more easily understood if they are horizontal, so we'll use `coord_flip` to make it horizontal.

```{r}
df%>%
  group_by(control)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE))%>%
  ggplot(aes(x=control,y=mean_debt,color=control))+
  geom_point(size=10)

```

*Quick Exercise: Create a dot plot showing average debt level by region*



## Conditional means using two predictors

We can use graphics to display conditional means at multiple levels of predictor levels. There are a couple of ways to get this done. When using bar plots we've got two basic tools: location and color. In the first example, we're going to plot debt as a function of control and the level of the college, 2 year or 4 year. We'll use position for control and color for the type of college. 


```{r}
df%>%
  group_by(control,preddeg)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE))%>%
  ggplot(aes(x=control,y=mean_debt,fill=preddeg))+
  geom_bar(stat="identity",position="dodge")


```


*Quick Exercise: Create either a bar plot or a dot plot showing debt levels by region and control*

## More Variables: faceting

We can continue this logic with three variables. Now we're going to summarize by control and degree and selectivity. Here we're going to use an additional tool in our arsenal: Faceting. Faceting means making multiple graphs with the same structure. In the code below, we will arrange positions based on control, color based on degree type, and faceting by selectivity. 


```{r}
df%>%
  group_by(control,preddeg,selective)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE))%>%
  ggplot(aes(x=control,y=mean_debt,fill=preddeg))+
  geom_bar(stat="identity",position="dodge")+
  facet_wrap(~selective)




```

_Side_ _Note_: Colors

What if we want to change colors? This is a little tricky for most people at first. `ggplot` thinks in terms of palettes, so you need to associate a palette with a characteristics of the graphic. Below, I replace the default palette with my own ugly one. 

```{r}

## Changing Colors
mypal<-c("lightblue","orange")



df%>%
  group_by(control,preddeg,selective)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE))%>%
  ggplot(aes(x=control,y=mean_debt,fill=preddeg))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_manual(values =mypal )+
  facet_wrap(~selective)




```

You can also use `RColorBrewer` which has a wide variety of palettes already built. Below I use the qualitative palette creatively named "Set1".

```{r}
df%>%
  group_by(control,preddeg,selective)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=TRUE))%>%
  ggplot(aes(x=control,y=mean_debt,fill=preddeg))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_brewer(palette="Set1" )+
  facet_wrap(~selective)

```

*Quick Exercise: Plot debt levels by *

## Graphs with many levels

If you have many levels it can be helpful to "flip" the graph to reflect that. Let's plot all 50 states by public and private institutions.

```{r}
df%>%
  group_by(stabbr)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=T))%>%
  ggplot(aes(x=stabbr,y=mean_debt))+
  geom_bar(stat="identity",position="dodge")
```

It plotted it, but it doesn't look very good! Let's try flipping the coordinates

```{r}
df%>%
  group_by(stabbr)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=T))%>%
  ggplot(aes(x=stabbr,y=mean_debt))+
  geom_bar(stat="identity",position="dodge")+
  coord_flip()
```

Better, but still awful. We need to do a couple of things. First, we should sort by the level of debt. 
To do that, we need to reorder the states in the order of debt levels. To do that, we need to work with the state variable `stabbr` as a factor.



This is fine, but it should really be in the order of the underlying variable. We can use `fct_reorder` to do this. 

_Side_ _Note_

What is a factor variable? In R, factor variables are used for categorical data. These are data elements that can take on one and only one value of a mutually exclusive and exhaustive list of elements.. We use factors when numbers won't work-- for characteristics like race or religion or political affiliation, or in this case, state. 

```{r}
df%>%
  group_by(stabbr)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=T))%>%
  mutate(stabbr=fct_reorder(stabbr,mean_debt))%>%
  ggplot(aes(x=stabbr,y=mean_debt))+
  geom_bar(stat="identity",position="dodge")+
  coord_flip()
  
```



Okay, better! Let's make the color scheme a function of the level of debt using `scale_fill_fermenter'.

```{r}
df%>%
  group_by(stabbr)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=T))%>%
  mutate(stabbr=fct_reorder(stabbr,mean_debt))%>%
  ggplot(aes(x=stabbr,y=mean_debt,fill=mean_debt))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_fermenter(palette="Set1")+
  coord_flip()
```


We can also use facets when there are multiple groups, so for instance if we want to work across regions:

```{r}
df%>%
  group_by(region,control,preddeg)%>%
  summarize(mean_debt=mean(grad_debt_mdn,na.rm=T))%>%
  ggplot(aes(x=control,y=mean_debt,fill=preddeg))+
  geom_bar(stat="identity",position="dodge")+
  facet_wrap(~region,nrow=2)

```


*No so quick exercise: Plot debt levels for public and private four-year institutions (use `filter`) by whether or not they're research institutions*

