---
title: "03-plot-means-inclass.Rmd"
author: "Doyle"
date: "2/11/2021"
output: github_document
---

```{r}
library(tidyverse)
library(RColorBrewer)
```

```{r}
df<-readRDS("sc_debt.Rds")
```


## Plotting Means: In Class Work

1. Using the student debt dataset, calculate and plot average SAT by region. Plot the result using a bar plot. Make it nicely labeled and beautiful in every way.

```{r}
df%>%
  group_by(region)%>%
  summarize(mean_sat=mean(sat_avg,na.rm=TRUE))%>%
  mutate(region=fct_reorder(region,mean_sat))%>%
  ggplot(aes(x=region,y=mean_sat,fill=region))+
  geom_bar(stat="identity",position="dodge")+
  theme_minimal()+
  theme(legend.position = "none")+
  xlab("Region")+ylab("Mean SAT Scores")

```


2. Repeat the above plot, but use a dot plot and reverse the axes. 


```{r}
df%>%
  group_by(region)%>%
  summarize(mean_sat=mean(sat_avg,na.rm=TRUE))%>%
  mutate(region=fct_reorder(region,mean_sat))%>%
  ggplot(aes(x=region,y=mean_sat,color=region))+
  geom_point(size=5)+
  theme_minimal()+
  theme(legend.position = "none")+
  xlab("Region")+ylab("Mean SAT Scores")+
  coord_flip()

```



3. Calculate and plot average SAT scores  by region  AND predominant degree type. Use a bar plot. 


```{r}
df%>%
  group_by(region,preddeg)%>%
  summarize(mean_sat=mean(sat_avg,na.rm=TRUE))%>%
  mutate(region=fct_reorder(region,mean_sat))%>%
  ggplot(aes(x=region,y=mean_sat,fill=preddeg))+
  geom_bar(stat="identity",position="dodge")+
  theme_minimal()+
  theme(legend.position = "none")+
  xlab("Region")+ylab("Mean SAT Scores")
```


4. Change the color scheme of the above plot. 

Using RColorBrewer
```{r}
df%>%
  group_by(region,preddeg)%>%
  summarize(mean_sat=mean(sat_avg,na.rm=TRUE))%>%
  mutate(region=fct_reorder(region,mean_sat))%>%
  ggplot(aes(x=region,y=mean_sat,fill=preddeg))+
  geom_bar(stat="identity",position="dodge")+
  theme_minimal()+
  xlab("Region")+ylab("Mean SAT Scores")+
  scale_fill_brewer(palette = "Set3")+
  labs(fill="Predominant Degree")
```



```{r}
mypal=c("lightblue","orange")

df%>%
  group_by(region,preddeg)%>%
  summarize(mean_sat=mean(sat_avg,na.rm=TRUE))%>%
  mutate(region=fct_reorder(region,mean_sat))%>%
  ggplot(aes(x=region,y=mean_sat,fill=preddeg))+
  geom_bar(stat="identity",position="dodge")+
  theme_minimal()+
  xlab("Region")+ylab("Mean SAT Scores")+
  scale_fill_manual(values=mypal)
```


5. Calculate and plot SAT scores by region and predominant degree with a separate plot (hint:facet_wrap) for public and private institutions.  
```{r}
df%>%
  group_by(region,preddeg,control)%>%
  summarize(mean_sat=mean(sat_avg,na.rm=TRUE))%>%
  mutate(region=fct_reorder(region,mean_sat))%>%
  ggplot(aes(x=region,y=mean_sat,fill=preddeg))+
  geom_bar(stat="identity",position="dodge")+
  theme_minimal()+
  theme(legend.position = "none")+
  xlab("Region")+ylab("Mean SAT Scores")+
  facet_wrap(~control)
```




