---
title: Working With Flat Data Files
output: github_document
---

Flat data is data that is arranged with one case per row, with one column per variable-- more or less. It's stored in a variety of formats, with different conventions. Our goal is to get it into the most useful format for analysis: what's known as tidy data. 

```{r} 
library(tidyverse)
library(haven)
library(readxl)
```


## CSV or other delimited files

We'll start with a csv file which is among the most common formats for datasets. CSV stands for _Comma _Separated _Value, meaning that each row is divided into cells by commas. An end of line completes the row. 


```{r}
#Delimited files

#Load in the HSB dataset from the UCLA statistical computing site

hsb<-read_csv(file="https://stats.idre.ucla.edu/wp-content/uploads/2016/02/hsb2-2.csv")

write_csv(hsb,path="hsb.csv")

#Check it out
head(hsb)

##Need these for later
my.names<-names(hsb)

#Write this in a variety of formats to be used later
write_delim(hsb, path="hsb.txt",delim="\t")

write_delim(hsb, path="hsb_semicolon.txt",delim=";")
gdata::write.fwf(data.frame(hsb),file="hsb.dat",sep="",colnames=FALSE)
```

*Quick exercise: write out the HSB file with a semicolon delimiter*


## Fixed width files

fixed width files are an older file format that you don't see as much of any more. To read these in, you need a file that tells you the locations of the different variables, known as column positions or locations. 

```{r}
## ------------------------------------------------------------------------

#Fixed width files

#You need to get the "widths" somehere, usually a data dictionary

my.widths=c(3,#id
            1, #female
            1, #race
            1, #ses
            1, #schtyp
            1, #prog
            2, #read
            2, #write
            2, #math 
            2, #science
            2 #socst
            )
            
my_positions<-fwf_widths(my.widths)

hsb3<-read_fwf("hsb.dat",
         col_positions  =my_positions)

head(hsb3)

names(hsb3)<-my.names

head(hsb3)

```

## Excel (sigh)

In most business applications you'll work with excel files the most often. To get these into shape you'll have to do some wrangling. Below I show how this is done with data in a common reporting format. 

```{r}
## ------------------------------------------------------------------------

#Excel files

##http://nces.ed.gov/programs/digest/d14/tables/dt14_204.10.asp

download.file("http://nces.ed.gov/programs/digest/d14/tables/xls/tabn204.10.xls",destfile="free.xls")

free<-read_excel("free.xls",skip=4,col_names=FALSE)
            
head(free)

# Now need to clean up 
#Get rid of unwanted columns

free2<-free[ ,-(c(3,6,9,12,15,18))]

#Get rid of unwanted rows
free2<-free2%>%filter(is.na(X1)==FALSE)

##50 states plus dc only
free2<-free2[2:52,]


head(free2)
tail(free2)

names(free2)<-c("state",
                "total.2000",
                "total.2010",
                "total.2011",
                "total.2012",
                "frl.2000",
                "frl.2010",
                "frl.2011",
                "frl.2012",
                "pc.frl.2000",
                "pc.frl.2010",
                "pc.frl.2011",
                "pc.frl.2012")
```

*Quick Exercise: Read in this file: http://nces.ed.gov/programs/digest/d14/tables/xls/tabn302.10.xls*

```{r}

download.file("http://nces.ed.gov/programs/digest/d14/tables/xls/tabn302.10.xls",destfile="grad.xls")

grad<-read_excel("grad.xls",skip=6,col_names=FALSE)

#Get rid of unwanted rows
grad<-grad%>%filter(is.na(X1)==FALSE)


```


## Tidy data

Tidy data follows two key principles: each column is one variable and one variable only, while each row is a case. Below, I show how to make the data from the above spreadsheet tidy, and why we would do this. 

```{r}

free_total<-free2%>%select(state,
                           total.2000,
                           total.2010,
                           total.2011,
                           total.2012)
names(free_total)<-c("state","2000","2010","2011","2012")

free_total<-free_total%>%gather(`2000`,`2010`,`2011`,`2012`,key=year,value=total_students)


frl_total<-free2%>%select(state,
                           frl.2000,
                           frl.2010,
                           frl.2011,
                           frl.2012)

names(frl_total)<-c("state","2000","2010","2011","2012")

frl_total<-frl_total%>%gather(`2000`,`2010`,`2011`,`2012`,key=year,value=frl_students)



free_tidy<-left_join(free_total,frl_total,by=c("state","year"))

free_tidy

## Total by year

free_tidy%>%group_by(year)%>%summarize(sum(frl_students))



pc.frl_total<-free2%>%select(state,
                           pc.frl.2000,
                           pc.frl.2010,
                           pc.frl.2011,
                           pc.frl.2012)

names(pc.frl_total)<-c("state","2000","2010","2011","2012")

pc_frl_total<-pc.frl_total%>%gather(`2000`,`2010`,`2011`,`2012`,key=year,value=pc_frl_students)

## ------------------------------------------------------------------------
```

*Quick Exericse: now add in percent of students eligible by state*

## Other programming languages

Other statistical programs have their own file formats. These are easy for these programs to read in. R can udnerstand all of them, if the `haven` packages is used. 

```{r}
#Other data files from stat programming language


# Stata

hsb_stata<-read_dta("https://stats.idre.ucla.edu/stat/stata/notes/hsb2.dta")

head(hsb_stata)


#SPSS
example_spss<-read_spss("https://stats.idre.ucla.edu/stat/data/binary.sav")

head(example_spss)

#SAS
hsb_sas<-read_sas("https://stats.idre.ucla.edu/wp-content/uploads/2016/02/hsb2.sas7bdat")

head(hsb_sas)

```

## Output

Most of the time, you should store your data as a csv file. This will ensure that pretty much anyone can take a look at it. If you're sure that the only users will be other R users (why would you be sure of this?), then feel free to save it as an `.Rdata` file. 

```{r}
## ------------------------------------------------------------------------
#Saving as an R file

save(free2,file="frl.Rdata")

#Outputting delimited

write_csv(free2,"frl.csv")
```
