# Scatterplots {#scatterplots}

Scatterplots are the best way to present data that has a continuous response variable. When creating scatterplots, the idea is to show ALL of the data, and then show how your model is summarizing the relationships in that data. 

## Motivation

Let's assume we have a grocery delivery service that we want to market to families that spend at least \$1,000  per quarter on groceries. We have a consumer datset that includes information about family income, number of children in the family and some other characteristics. We want to use this to isolate families that are in the desired spending range. 

## Setup
The code for today starts with the normal set of preliminaries, opening up the `cex.RData` dataset and creating a codebook. 

```{r,echo=FALSE}

library(tidyverse)
library(forcats)

load("cex.RData")

my.vals<-names(cex)
explanation<-c("id",
               "education: 10 8th grade, 11 9-12, 12 hs grad,13 some college, 14 assoc, 15 bach 16 masters, 17 prof",
"education of second person",
"urban: 1=urban, 2=not",
"race: 1 white, 2 black, 3 native american, 4 asian 5 pacific islander 6 multi",
"race of other person",
"income class: 1: 0-10k, 2 5-10k, 3 10-15k 4 15 to 20k 5 20-30k 6 30-40k 7 40-50k 8 50-70k 9 70k+",
"inc_rank: percentile rank of income",
"sex of reference: 1 male 2 female",
"sex of other person",
"reference person hispanic? 1 hispanic 2 non",
"other person hispanic? 1 hispanic 2 non",
"poverty: 1 below poverty, 2 not",
"region 1 Ne 2 MW 3 South 4 West",
"family size: ",
"family type 1 hw, 2 hw own children, oldest under 6, 3 hw own children oldest 6-17, 4 hw own children oldest over 17, 5 all other hws, 6 one parent male, kids under 18, 7 one parent female, kids under 18, 8 single, 9 other ",
"Children's ages: 0 no kids, 1 all less 6 2 oldest bt 6 and 11, 3 All children between 6 and 11
 4 Oldest child between 12 and 17 and at least one child less than 12
 5 All children between 12 and 17
 6 Oldest child greater than 17 and at least one child less than 17
 7 All children greater than 17 one under 6",
"quarter and year",
"Dining out",
"Grocery",
"Grocery non-food",
"Grocery food",
"Alcohol at home",
"Alcohol out",
"Other store expenditures",
"Cigarrettes",
"Transportation to work"
)

codebook<-data.frame(my.vals, explanation)

#View(codebook)

## Quick transformation of income rank

cex<-cex%>%mutate(inc_rank=inc_rank*100)
```

*Side Note: Factors*

Factors in R denote variables that encompass a set of mutually exclusive and exhaustive categories. When a factor is included in a regression, R will encode it as a series of dummy (binary) variables, one for each level of the factor. The coefficient for each level of the factor is relative to the reference (omitted) category. Let's run a regression of grocery food spending on age of children and see what we get.

```{r}

## Get levels of childage right

cex<-cex%>%mutate(childage=fct_recode(f=childage,
                     "No kids"="0", 
"All < 6"="1",
"Oldest bt 6 and 11,one <6"="2",
"All b/t 6 and 11"="3",
 "Oldest b/t 12 and 17 & one child < 12"="4",
 "All b/t 12 and 17"="5",
 "Oldest > 17 and one < 17"="6",
 "All > 17 "="7"
))

with(cex, prop.table(table(childage)))

#Conditional Means
cex%>%group_by(childage)%>%summarize(mean_groc=mean(grocery_food,na.tm=TRUE))

#Regression
factor_mod<-lm(grocery_food~childage,data=cex);summary(factor_mod)

#Predictions from model
print(factor_mod$coefficients[2:8]+factor_mod$coefficients[1])
```

The reference for the category (the excluded category) is families with no children. Each coefficient is in comparison to this category, indicating how much more or less families with children in each age group spend than families with no children on food groceries. Notice that in this case with no additional covariates, the intercept plus the coefficient for each group is exactly equal to the conditional mean for that group.  

## Bivariate Regression

We begin with a simple model of grocery food spending as a function of income rank. As we found out last week, grocery food spending has a long right tail, and is a good candidate for a log transformation. We'll be using that log transformation throughout. 

Our first step should be to plot the data. Today, we'll be using the `ggplot2` library, which is a highly functional implementation of what's known as the grammar of graphics. In a very small nutshell, the grammar of graphics refers to laying out a graphic in a series of layers. For our first scatterplot, we first specify the data that we'll be drawing on, then the "aesthetic" of the graphic, which will be based on our x and y variables from our regression. I then specify the first layer, which is a series of points defined by the intersection of the x and y variables. 


```{r}
#Plot Spending vs. Income
g1<-ggplot(data=cex,
           aes(x=inc_rank,y=(grocery_food+1))
           )

g1<-g1+geom_point() # Add points at x and y
g1
```

This first graphic shows the basic problem with the data, which is the highly non-normal shape of the outcome variable. No matter! We'll keep moving forward. The next line of code adds a layer with a regression line to the plot. 

```{r}
g1<-g1+geom_smooth(method="lm")
g1<-g1+ylab("Grocery Spending")+xlab("Income Rank")
g1
```

It's also really hard to see. We can use conditional means to help out with that problem. Let's get the average amount of grocery spending at every percentile level of `inc_rank`.  Notice the use of `round` to get income percentiles that are at two digits only. 

```{r}

cex_sum<-cex%>%mutate(inc_rank_r=round(inc_rank))%>%group_by(inc_rank_r)%>%summarize(groc_mean=mean(grocery))

g1a<-ggplot(cex_sum,aes(x=inc_rank_r,y=groc_mean))

g1a<-g1a+geom_point()

g1a<-g1a+ylab("Grocery Spending")+xlab("Income Rank")

g1a

```


Better! Simplifying data can help. 

We can add a regression line to this simpler data

```{r}
g1b<-g1a+geom_smooth(method=lm) # Add a line
g1b
```

```{r}
g1c<-g1a+geom_smooth(method=loess) # Add a line
g1c
```


As we figured out in class last week, a simple line plot doesn't really do it. We need to transform that dependent variable. Luckily, ggplot has this kind of transformation built in. 

```{r}
g1<-ggplot(data=cex,
           aes(x=inc_rank,y=(grocery_food+1))
           )
# Set nice looking breaks, the ones used by ggplot for log transform typically don't work
my.breaks=c(0,100,500,1000,2500,5000,10000,20000)
#Change the scale 
g1<-g1+scale_y_continuous(trans="log",breaks=my.breaks)
g1<-g1+geom_point() # Add points
g1<-g1+geom_smooth(method=lm)  #Add a Line
g1<-g1+ylab("Grocery Spending")+xlab("Income Percentile (0-100)") #Nice labels
g1
```

Notice how different the steps are on the y axis now. But this shows that our line actually fits the data much better once we work on the log scale. We're now ready to go ahead and run our first model based on what we've learned. 

```{r}
#First model

m1<-lm(log(grocery_food+1)~inc_rank,data=cex);summary(m1)

```

_Quick Exercise_ Create a similar graphic, but this time use other store expenditures as the dependent variable. Run a regression with store expenditures as the dv and income rank as the only independent variable. 

## Multiple Regression

The next step is to add covariates. I'll be working with the variable `childage` which is a factor that summarizes the ages of kids in the family. I'm going to set the color of the markers by the `childage` factor. 


```{r}
g2<-ggplot(data=cex,
            aes(x=inc_rank,y=(grocery_food+1),
                color=childage #notice the color option
                ))

g2<-g2+scale_y_continuous(trans="log",breaks=my.breaks)
g2<-g2+geom_point(size=.25)
g2<-g2+theme(legend.position="bottom"  , legend.title = element_blank())
g2<-g2+ylab("Grocery Spending")+xlab("Income Rank")
g2

```

Our first graphic is a bit complex, but shows that those in the "0" category, with no kids, have consistently lower spending on groceries across income levels. So let's drop that group and see what the graphic looks like for those with at least one kid of any age. 

```{r}
g2a<-ggplot(data=filter(cex,as.character(cex$childage)!="No kids"),
            aes(x=inc_rank,y=(grocery_food+1),color=childage)
           )
g2a<-g2a+scale_y_continuous(trans="log",breaks=my.breaks,limits=c(100,8000)) #Changed overall scale
g2a<-g2a+geom_point(size=.25) 
g2a<-g2a+scale_colour_hue(l=50) #Darker palette
g2a<-g2a+theme(legend.position = "bottom" , legend.title = element_blank())
g2a<-g2a+ylab("Grocery Spending")+xlab("Income Rank")
g2a
```

Notice the higher spending levels at every income level among those with children between 6 and 17. 

```{r}

#Model 2: with kids

mod2<-lm(log(grocery_food+1)~inc_rank+childage, data=cex,na.action=na.exclude);summary(mod2)

```


Now let's take a look at this model plotted against the actual data. I'm going to use the `alpha` setting to make the dots more transparent. I'm also going to make the dots smaller via the size specification. 
```{r}

cex<-cex%>%mutate(mod2_pred=exp(fitted(mod2)))

g3<-ggplot(cex,aes(x=inc_rank,y=grocery_food))
g3<-g3+geom_point(alpha=.2,size=.25)
g3<-g3+scale_y_continuous(trans="log",breaks=my.breaks)
g3<-g3+geom_smooth(data=cex,(aes(x=inc_rank,y=mod2_pred)))
g3<-g3+xlab("Income Percentile (0-100")+ylab("Grocery Food Spending")
g3
```


## Plotting more complex models

As we add more variables to the model, it can get more difficult to plot relationships. One very good option is to plot lines based on a hypothetical set of data. Below, I create a hypothetical set of data that include values of income across the range of income, and includes values for every level of `childage`.


```{r}
#Prediction, then plotting

hypo.data<-data.frame(expand.grid(
          #range of income
          inc_rank=seq(min(cex$inc_rank,na.rm=TRUE),
                       max(cex$inc_rank,na.rm=TRUE),
                       length=100), 
          #All levels of childage
          childage=levels(cex$childage)                                     
                                      )
                          )
head(hypo.data)
```

Now, using my estimates from model 2, I predict what would happen to these hypothetical individuals.  Once I've got my prediction, I transform it back out of the log scale into the "response" level of dollars. 

```{r}
#Predict using mod2 on hypothetical data
hypo.pred<-predict(mod2,
                   newdata=hypo.data,
                   interval="prediction",
                    se.fit=TRUE
                   )

hypo.data<-data.frame(hypo.data,hypo.pred$fit) #Add fit to the dataset
hypo.data$groc.p<-exp(hypo.data$fit) #exponentiate fit
hypo.data$groc.lwr<-exp(hypo.data$lwr) #exponentiate lower boundary
hypo.data$groc.upr<-exp(hypo.data$upr) #exponentiate upper boundary
head(hypo.data)
```


Now we can plot the result, using the `geom_smooth` layer to give us lines for every level of `childage`. 


```{r}
g4<-ggplot(data=hypo.data,aes(x=inc_rank,y=groc.p,color=fct_reorder(f=childage,-groc.p))) #notice color
g4<-g4+geom_smooth(method=lm,se=FALSE)
g4<-g4+theme(legend.position="bottom",legend.title = element_blank())
g4<-g4+xlab("Income Rank")+ylab("Predicted Grocery Spending")
g4
```

*Question: Based on our motivating example, where should we be advertising?*

To show this in the data we can break it out for every type of child age grouping. 

```{r}
## Resort child age for graphic

cex<-cex%>%mutate(childage=fct_reorder(f=childage,-grocery_food))

g5<-ggplot(cex,aes(x=inc_rank,y=grocery_food,color=childage))
g5<-g5+geom_point(alpha=.5)
g5<-g5+geom_smooth(method="lm",color="black")
g5<-g5+facet_wrap(~childage,ncol=3)
g5<-g5+scale_y_continuous(trans="log",breaks=my.breaks)
g5<-g5+xlab("Income Rank, 0-100")+ylab("Grocery Spending")
g5<-g5+theme(legend.position="none") #Suppress legend, not needed

g5

```

_(Not so) Quick Exercise_ Run a different model, this time using `fam_type` as a factor variable. Plot the estimated relationship between income rank and food spending for each family type. 
