---
title: "Plots for Classifiers"
author: "Will Doyle"
date: "August 7, 2018"
output: html_document
---

Plots are rarely used in the context of classification, but they can aid understanding. I'll show three ways of thinking about plots for classification: bar graphs, heatmaps, and plotting the probability predictions from a logit model.


We'll continue working with the random acts of pizza dataset. 

```{r}
library(tidyverse)
library(modelr)
library(knitr)
load("za.RData")
```

We always want to start with a cross tab of our dependent variable as a function of other variables. We structure cross tabs in a very particular way for the purposes of research: the independent variable goes on the rows, and the dependent variable goes on the columns. If proportions or percentages are going to be calculated, they should be calculated across rows. 

Recalling our previous lesson, let's look at a crosstab of `got_pizza` with the independent variable of `student`

```{r}
tab_student<-with(za,table(student,got_pizza))
```

If we want to make this a little better, we can change the row and column titles

```{r}
colnames(tab_student)<-c("No Pizza","Received a Pizza")
kable(tab_student)
```

If we want to add proportions to this table, we can it like so:
```{r}
tab_student_prop<-prop.table(tab_student,margin=1)
kable(tab_student_prop)
```

Sometimes  (okay, all the times) audiences prefer percentages. Easy enough to do:
```{r}
kable(round(tab_student_prop*100,2))
```

If you want to include a third variable in a cross tab, that requires splitting the dataset. For instance, if we want to know the proportion of posts that include "student" AND "grateful" that received pizza, we would do this:

```{r}
tab_student_grateful<-with(filter(za,
                           as.character(grateful)=="Grateful in post"),
                           table(student,got_pizza))
```

*Outcome by "Student" AND "Grateful"*
```{r}
prop.table(tab_student_grateful,margin=1)%>%kable()
```


## Barcharts

To format the data for barcharts, we make use of the concept of conditional means. Let's use two variables to calculate the probability of receiving a pizza: poor and student. 

```{r}
za_sum<-za%>%
  group_by(poor,student)%>%
  summarize(prob_pizza=mean(got_pizza,na.rm=TRUE))

gg1<-ggplot(za_sum,aes(y=prob_pizza,x=student,fill=poor))
gg1<-gg1+geom_bar(stat="identity",position="dodge")
gg1<-gg1+xlab("Student Status")+ylab("Pr(Pizza)")
gg1<-gg1+theme(legend.title=element_blank())
gg1<-gg1+geom_text(aes(label=round(prob_pizza,2)),
                   position=position_dodge(width=.9),
                   vjust=-.25)
gg1
```


## Heat Maps

```{r}
za<-za%>%mutate(score_decile=ntile(score,5),
                karma_decile=ntile(karma,5))


za_sum<-za%>%group_by(score_decile,karma_decile)%>%
  summarize(prob_pizza=mean(got_pizza,na.rm=TRUE))%>%
  arrange(-prob_pizza)

za_sum<-za_sum%>%filter(!(is.na(score_decile)),!(is.na(karma_decile)))

gg<-ggplot(za_sum,
           aes(x=as.factor(score_decile),
               y=as.factor(karma_decile),fill=prob_pizza))
gg<-gg+geom_tile()
gg<-gg+scale_fill_gradient(low="white",high="red")
gg<-gg+xlab("Emotional Valence")+ylab("Reddit Karma")
gg<-gg+theme(legend.title=element_blank())
gg

```


## Plotting by probabilities from models
```{r}

logit_mod<-glm(got_pizza~
             karma+
             total_posts+
             raop_posts+
             student+
             grateful,
             data=za,
            na.action=na.exclude,
            family=binomial(link="logit"),
               y=TRUE)


```


```{r}
hypo_data<-data_grid(za, 
                     total_posts=seq_range(total_posts,n=100),
                                           karma=mean(karma,na.rm=TRUE),
                                           raop_posts=mean(raop_posts,na.rm=TRUE),
                     grateful=levels(grateful)[1],
                                           student=levels(student))%>%
              mutate(pred=predict(logit_mod,newdata=.,type="response"))

gg<-ggplot(hypo_data,
           aes(x=total_posts,y=pred,color=student))
gg<-gg+geom_line()
gg<-gg+xlab("Total Posts")+ylab("Predicted Probability of Pizza")
gg
```

