---
title: "10-cross-validation"
author: "Will Doyle"
date: "4/20/2021"
output: html_document
---

```{r}
library(tidyverse)
library(tidymodels)
```

```{r}
load("at.Rdata")
```

## Cross Validation: In Class Work

1. Fit a linear model predicting monthly wages using the attrition dataset (`at.Rdata`) and cross validate those results. Competition: lowest average RMSE wins, with bonus points for low sd.

```{r}
lm_formula<-as.formula("MonthlyIncome~
                       Education+
                       EducationField+
                       Age+
                       Department")

```


```{r}
at%>%
  ggplot(aes(x=MonthlyIncome))+
  geom_density()
```


```{r}

lm_rec <- recipe(lm_formula, data = at) %>%
  step_log(all_outcomes())%>%
  step_zv(all_numeric()) %>% # drop any zero variance
  step_naomit(all_predictors())%>% ## drop any missing data
  step_dummy(all_nominal()) 

```


## Set the kind of model to run
```{r}
lm_fit <- 
  linear_reg() %>% 
  set_engine("lm")
```


## Specify the resampling: bootstrap resampling
```{r}
boot <- bootstraps(at,times = 100) ##1000 is usual minimum
```


## Fit Bootstrap Reasmpling
```{r}
lm_boot<-
  lm_fit%>%
  fit_resamples(
    lm_rec, ## Recipe: preps the data
    boot, ##resampling plan
    control=control_resamples(save_pred = TRUE)
  )
```

```{r}
lm_boot%>%
  collect_metrics()
```

```{r}
lm_boot%>%
  unnest(.metrics)%>%
  filter(.metric=="rmse")%>%
  ggplot(aes(x=.estimate))+
  geom_density()
```


2. Fit a logistic regression predicting attrition using the attrition dataset (`at.Rdata`) and cross validate those results. Competition: highest average AUC wins, with bonus points for low sd. 


```{r}
at<-at%>%
  mutate(Attrition=as_factor(Attrition))
```


```{r}

logit_formula<-as.formula("Attrition~
                          MonthlyIncome+
                         Education+
                       EducationField+
                       Age+
                       Department")

```


```{r}

logit_rec <- recipe(logit_formula, data = at) %>%
  step_zv(all_numeric()) %>% # drop any zero variance
  step_naomit(all_predictors())%>% ## drop any missing data
  step_dummy(all_nominal(),-all_outcomes()) 

```

```{r}
logit_rec%>%
  prep()
```



```{r}
logit_boot <- bootstraps(at,times = 100) ##1000 is usual minimum
```


## Specify Model
```{r}
logit_fit<-
  logistic_reg(mode="classification")%>%
  set_engine("glm")
```


```{r}
logit_boot <- 
  logit_fit%>%
  fit_resamples(
    logit_rec,
    logit_boot,
    metrics = metric_set(roc_auc, sens, spec)
  )
```




```{r}
logit_boot%>%
  collect_metrics()
```


## Plot distribution of AUC
```{r}
logit_boot%>%
  unnest(.metrics)%>%
  filter(.metric=="roc_auc")%>%
  ggplot(aes(x=.estimate))+
  geom_density()
```



