---
title: "10-cross-validation"
author: "Will Doyle"
date: "4/20/2021"
output: html_document
---

```{r}
library(tidyverse)
library(tidymodels)
```

```{r}
load("at.Rdata")
```

## Cross Validation: In Class Work

1. Fit a linear model predicting monthly wages using the attrition dataset (`at.Rdata`) and cross validate those results. Competition: lowest average RMSE wins, with bonus points for low sd.

```{r}

```


## Set the kind of model to run
```{r}
lm_fit <- 
  linear_reg() %>% 
  set_engine("lm")
```


## Specify the resampling: bootstrap resampling
```{r}
boot <- bootstraps(pd,times = 100) ##1000 is usual minimum
```


## Fit Bootstrap Reasmpling
```{r}
lm_boot<-
  lm_fit%>%
  fit_resamples(
    lm_rec, ## Recipe: preps the data
    boot, ##resampling plan
    control=control_resamples(save_pred = TRUE)
  )
```


2. Fit a logistic regression predicting attrition using the attrition dataset (`at.Rdata`) and cross validate those results. Competition: highest average AUC wins, with bonus points for low sd. 


## Specify Model
```{r}
logit_fit<-
  logistic_reg(mode="classification")%>%
  set_engine("glm")
```



