---
title: "Clustering-- in class"
author: "Doyle"
date: "11/21/2019"
output: github_document
---

```{r}
library(tidyverse)
library(purrr)
library(LICORS)
library(factoextra)
```


For this in class work, complete the following steps:

1. Subset the data for 1995,2000,2005,2010 and 2015


```{r}
load("nba.Rdata")
```

```{r}
nba_df_all%>%
  filter(minutes_played>500)%>%
  group_by(player,year)%>%
  top_n(1,minutes_played)%>%
  ungroup()%>%
   select(player,
         pts,
         minutes_played,
         fg,
         fg_attempts,
         fg_percent,
         three_pointers,
         three_point_attempts,
         three_point_percent,
         two_pointers,
         two_point_attempts,
         two_point_percent,
         effective_fg_percent,
         free_throws,
         free_throw_attempts,
         free_throw_percent,
         off_rebound,
         def_rebound,
         total_rebound,
         assists,
         steals,
         blocks,
         turnovers,
         fouls,
         year)%>%
               drop_na()->nba_df_sub

player_id<-nba_df_sub$player
points<-nba_df_sub$pts
year<-nba_df_sub$year

# Scale the dataset
nba_df_sub%>%
  select(-player,-pts,-year)%>%
  mutate_all(scale)->nba_df_cluster_all

nba_df_cluster_all$year<-year
```

1. Repeat the steps in K-means clustering from Tuesday for each year.

This generates the elbow plots. Notice that it begins by splitting the overall dataset by year,
then mapping that to the function using map(~). The data frame is implied as ., and the `select` operator drops the year, which we don't want in the analysis. The output is saved as a data frame using `as_tibble`. The elements of the data frame can then be called. 
```{r}
nba_df_cluster_all%>%
split(.$year)%>%  
map(~fviz_nbclust(select(.,-year),
             FUNcluster=kmeanspp,
             method="wss"))->elbow_plots

elbow_plots$`1995`
elbow_plots$`2000`
elbow_plots$`2005`
elbow_plots$`2010`
elbow_plots$`2015`

```


This runs the kmeans clustering. I'm using the same basic approach as above. This time the output is a dataframe of all of the results of the kmeans. The first row of that data frame is the cluster assignments. When "unnested" these form a vector of cluster assignments that can be added to the original dataset. 
```{r}
nba_df_cluster_all%>%
  mutate_all(.funs=as.numeric)%>% # Make sure everything is numeric
  split(.$year)%>% # Split by year
  map(~kmeanspp(data=select(.,-year),k=5,n=100,nstart=50))%>% # Apply kmeans clustering to each year
  as_tibble()->cluster_tbl ## output results as data frame

## Clusters are in row 1, unlist and add back to analysis dataset 

nba_df_cluster_all$cluster<-unlist(cluster_tbl[1,])
```



1. Try to describe the groups you identify for each of the five years.


Again, same basic setup. This time though, I'm going to turn the plotting code chunk into a function, then apply that function to each element of the list created by splitting the data frame by year. Again writing the results using `as_tibble` I get a data frame of plots, which can then be called.

```{r}
plot_cluster<-function(x){
  ## plot function: x is a data frame, that includes a cluster variable
  x%>%
  group_by(cluster)%>%
  summarize_all(.funs=mean)%>%
  pivot_longer(cols=(-cluster),
               names_to="stat",
               values_to = "mean_results")->clus_results

gg<-ggplot(clus_results,aes(x=as_factor(stat),
                            y=mean_results,
                            color=stat))
gg<-gg+geom_point()
gg<-gg+facet_wrap(~cluster,ncol=1)
gg<-gg+geom_hline(yintercept=0)
gg<-gg+coord_flip()
gg<-gg+theme(legend.position = "none")
gg<-gg+theme(axis.text.x=element_text(size=.75))
gg
}

nba_df_cluster_all%>%
  split(.$year)%>%
  map(~plot_cluster(select(.,-year)))%>%
  as_tibble()->plots


plots$`1995`
plots$`2000`
plots$`2005`
plots$`2010`
plots$`2015`
```



1. Generate a predictive analysis predicting points scored based only on the groups identified for each year.

And one more time, this time 
```{r}
nba_df_cluster_all$pts<-points

nba_df_cluster_all%>%
  split(.$year)%>%
  map(~lm(pts~as.factor(cluster),data=select(.,-year)))%>%
  as_tibble()->reg_results
  
summary(reg_results$`1995`)
summary(reg_results$`2000`)
summary(reg_results$`2005`)
summary(reg_results$`2010`)
summary(reg_results$`2015`)
```

