---
title: "Untitled"
author: "Doyle"
date: "9/12/2019"
output: html_document
---

1. Using the attrition dataset, calculate and plot the probability of attrtion by education. Plot the result using a bar plot. Make it nicely labeled and beautiful in every way.

```{r}
at_sum<-at%>%group_by(Education)%>%
  summarize(mean_attr=mean(attrit))

gg<-ggplot(at_sum,aes(x=reorder(Education,mean_attr),y=mean_attr,fill=Education))   #,fill=Education))
gg<-gg+geom_bar(stat="identity")
gg

```

```{r}

at_sum<-at%>%
  group_by(Education,JobSatisfaction)%>%
  summarize(mean_attr=mean(attrit))%>%
  ungroup()

at_sum<-at_sum%>%
  mutate(Education=as_factor(Education))%>%
  mutate(JobSatisfaction=as_factor(JobSatisfaction))

gg<-ggplot(at_sum,aes(x=Education,y=mean_attr,fill=JobSatisfaction))
gg<-gg+geom_bar(stat="identity",position="dodge")
gg


```



2. Repeat the above plot, but use a dot plot and reverse the axes. 
3. Calculate and plot the probability of attrition by education AND the employee's job satisfaction. Use a bar plot. 
4. Change the color scheme of the above plot. 
5. Calcualte and plot the probability of attrition by job satisfaction AND work life balance, with a separate plot (hint:facet_wrap) for every level of education. 