---
title: "Assignment 1 Key"
author: "Will Doyle"
date: "September 3, 2019"
output: github_document
---

*1. A sentence that says "Hello, World"*

Hello, world!

<OR>

```{r}
print("Hello, World!")
```


*2. R output that summarizes one of the variables in the `colllege.Rdata` dataset*

```{r}
library(tidyverse)

# I left the college.Rdata in the lessons folder, so I will find it there
load("../lessons/college.Rdata")

sc%>%summarize(mean_cost=mean(costt4_a,na.rm=TRUE))

```


*3. R output that shows a scatterplot for two of the variables in the `college.Rdata` dataset.*

```{r}
gg<-ggplot(data=sc,aes(x=costt4_a,y=debt_mdn))
gg<-gg+geom_point()
gg<-gg+geom_smooth()
gg
```


## Stretch items


*1. Calculate the average earnings for individuals at the most selective colleges, then compare that with individuals at the least selective colleges in the dataset. *


```{r}
sc%>%filter(adm_rate<.1)%>%summarize(mean_earnings=mean(md_earn_wne_p6,na.rm=TRUE))

sc%>%filter(adm_rate>.3)%>%summarize(mean_earnings=mean(md_earn_wne_p6,na.rm=TRUE))

```


*2. Find a way to determine whether colleges with very high SAT scores tend to be larger or smaller than colleges with low SAT scores. *

```{r}
sc%>%filter(sat_avg>1400)%>%summarize(mean_enroll=mean(ugds, na.rm=TRUE))
sc%>%filter(sat_avg<1000)%>%summarize(mean_enroll=mean(ugds, na.rm=TRUE))
```


*3. Plot the relationship between cost and debt. What do you see? Does this surprise you?*

```{r}


gg<-ggplot(data=sc,aes(x=costt4_a,y=debt_mdn))
gg<-gg+geom_point()
gg<-gg+geom_smooth()
gg
```
*4. Now, provide separate plots for cost and debt by control of the institution. *

```{r}

gg<-ggplot(data=(sc%>%filter(control==1)),aes(x=costt4_a,y=debt_mdn))
gg<-gg+geom_point()
gg<-gg+geom_smooth()
gg


gg<-ggplot(data=(sc%>%filter(control==2)),aes(x=costt4_a,y=debt_mdn))
gg<-gg+geom_point()
gg<-gg+geom_smooth()
gg
```