---
title: "Assignment 2 Key"
output: html_document
author: "Will Doyle"
---


## Load packages
```{r}
library(tidyverse)
library(ModelMetrics)
```


## Load Data
```{r}
load("../lessons/pd.RData")
```


1. Calculate the mean of the outcome.
```{r}
pd%>%summarize(mean(homeown_rate))
```



2. Use your mean as a prediction: Create a new variable that consists of the mean of the outcome.

```{r}
pd<-pd%>%mutate(mean_homeown_rate=mean(homeown_rate))
```


3. Calculate a summary measure of the errors for each observation---the difference between your prediction and the outcome. 


```{r}
rmse_one<-rmse(pd$mean_homeown_rate,pd$homeown_rate) ; rmse_one
```

4. Calculate the mean of the outcome at levels of a predictor variable.  

```{r}
pd<-pd%>%mutate(pop65_level=ntile(pop65p,4))

pd%>%group_by(pop65_level)%>%
  summarize(homeown_rate_pop65=mean(homeown_rate))
```


5. Use these conditional means as a prediction: for every county, use the conditional mean to provide a ''best guess'' as to that county's level of the outcome.  

```{r}
pd<-pd%>%group_by(pop65_level)%>%
  mutate(homeown_rate_pop65=mean(homeown_rate))
```


6. Calculate a summary measure or two of the error in your predictions.

```{r}

rmse_two<-rmse(pd$homeown_rate_pop65,pd$homeown_rate); rmse_two
```

7. Repeat the above process using the tool of conditional means, try to find 3-4 variables that predict the outcome with better (closer to 0) summary measures of error. Report the summary measures of error and the variables (as text in your `.Rmd` file).

```{r}
pd<-pd%>%mutate(inc_level=ntile(per_capita_inc,4))%>%
  mutate(coll_level=ntile(coll_grad_pc,4))%>%
           mutate(home_value_level=ntile(median_home_val,4))


pd<-pd%>%group_by(pop65_level,inc_level,coll_level,home_value_level)%>%
  mutate(homeown_rate_4=mean(homeown_rate))

rmse_three<-rmse(pd$homeown_rate_4,pd$homeown_rate);rmse_three
```

The root mean squared errors are respectively, `r round(rmse_one,2)`, `r round(rmse_two,2)` and `r round(rmse_three,2)`. 

