---
title: "02-assignment"
author: "Doyle"
date: "9/10/2019"
output: html_document
---


```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(gridExtra)
library(ModelMetrics)
```

```{r}
load("../lessons/pd.RData")
```


1.  Calculate the mean of the outcome.

```{r}
pd%>%summarize(mean_homeown=mean(homeown_rate,na.rm=TRUE))
```



2.  Use your mean as a prediction: Create a new variable that consists    of the mean of the outcome.


```{r}
pd%>%
  mutate(mean_homeown=mean(homeown_rate,na.rm=TRUE))->pd
```

3.  Calculate a summary measure of the errors for each observation—the
    difference between your prediction and the outcome.
    
```{r}
rmse_uncond<-rmse(pd$homeown_rate,pd$mean_homeown);rmse_uncond
```
    
4.  Calculate the mean of the outcome at levels of a predictor
    variable.  
    
```{r}
pd%>%
  mutate(coll_grad_level=ntile(coll_grad_pc,4))%>%
  group_by(coll_grad_level)%>%
  summarize(mean_homeown_coll=mean(homeown_rate,na.rm=TRUE))
```
    
    
5.  Use these conditional means as a prediction: for every county, use
    the conditional mean to provide a ‘’best guess’’ as to that county’s
    level of the outcome.


```{r}
pd%>%
  mutate(coll_grad_level=ntile(coll_grad_pc,4))%>%
  group_by(coll_grad_level)%>%
  mutate(mean_homeown_coll=mean(homeown_rate,na.rm=TRUE))->pd
```
        
    
6.  Calculate a summary measure of the error in your predictions.

```{r}
rmse_one<-rmse(pd$homeown_rate,pd$mean_homeown_coll)
rmse_one
```


7.  Repeat the above process using the tool of conditional means, try to
    find 3-4 variables that predict the outcome with better (closer to
    0) summary measures of error. Report the summary measures of error
    and the variables (as text in your `.Rmd` file).
    
```{r}
pd%>%
  mutate(percap_level=ntile(percapinc.2010,4))%>%
  group_by(coll_grad_level,percap_level)%>%
  mutate(mean_homeown_2=mean(homeown_rate,na.rm=TRUE))->pd
```

```{r}
rmse_two<-rmse(pd$homeown_rate,pd$mean_homeown_2)
rmse_two
```
    
    
