---
title: "Assignment 3 Key"
output: github_document
---

Use the attrition dataset to answer the following questions. As always, your assignment must be turned in as a .Rmd file. 

```{r}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(forcats)
library(RColorBrewer)

load("../lessons/at.Rdata")
```


1. Create a graph that shows the distribution of monthly income.
```{r}
gg<-ggplot(data=at,aes(MonthlyIncome))
gg<-gg+geom_histogram(bins=50)
gg
```



2. Create a graph that shows the average level of monthly income by field of education.

```{r}
at_sum=at%>%
  group_by(Education)%>%
  summarize(mean_salary=mean(MonthlyIncome))

gg<-ggplot(data=at_sum, aes(x=as.factor(Education),y=mean_salary,fill=as.factor(Education)))
gg<-gg+geom_bar(stat="identity",position="dodge")
gg<-gg+xlab("Education Level")+ylab("Average Salary")
gg<-gg+theme(legend.position="none")
gg
```
3. Create another graph that shows average level of monthly income by field of education and gender.

```{r}
at_sum=at%>%
  group_by(Education,Gender)%>%
  summarize(mean_salary=mean(MonthlyIncome))

gg<-ggplot(data=at_sum, aes(x=as.factor(Education),y=mean_salary,fill=as.factor(Gender)))
gg<-gg+geom_bar(stat="identity",position="dodge")
gg<-gg+xlab("Education Level")+ylab("Average Salary")
gg<-gg+theme(legend.title=element_blank())
gg
```



4. Create a graph that shows average levels of monthly income by field of education, gender and job level (scale of 1-5, highest ranked employees are 5)



```{r}
at_sum=at%>%
  group_by(Education,Gender,JobLevel)%>%
  summarize(mean_salary=mean(MonthlyIncome))

gg<-ggplot(data=at_sum, aes(x=as.factor(Education),y=mean_salary,fill=as.factor(Gender)))
gg<-gg+geom_bar(stat="identity",position="dodge")
gg<-gg+xlab("Education Level")+ylab("Average Salary")
gg<-gg+facet_wrap(~JobLevel)
gg<-gg+theme(legend.title=element_blank())
gg
```

5. Create a graph that shows average levels of monthly income by field of education, gender, job level and department. 

```{r,fig.height=9}

at_sum<-at%>%
  group_by(Education,Gender,JobLevel,Department)%>%
  summarize(salary_avg=mean(MonthlyIncome))%>%
  ungroup()%>%
  arrange(salary_avg)

at_sum

## One Solution: a new variable for every type of employee defined by travel, gender, marital status, and happiness with work/life

at_sum$grouping<-paste0("Ed Level: ",
                       at_sum$Education,
                       ", ",
                       at_sum$Gender,
                       ", ",
                       "Job Level: ",
                       at_sum$JobLevel,
                       ", ",
                       at_sum$Department)

at_sum$grouping<-as.factor(at_sum$grouping)

## Showing only highly paid employees
at_sum<-filter(at_sum, salary_avg>4000)

at_sum%>%select(grouping,salary_avg)

gg<-ggplot(at_sum,aes(x=fct_reorder(grouping,salary_avg),y=salary_avg))
gg<-gg+geom_bar(stat="identity",aes(fill=Department))
gg<-gg+coord_flip()
gg<-gg+xlab("")+ylab("Salary")
gg

#OR#

gg<-ggplot(at_sum,aes(x=as.factor(Education),y=salary_avg))
gg<-gg+geom_bar(stat="identity",position="dodge",aes(fill=Department))
gg<-gg+facet_wrap(~Gender+JobLevel)
gg<-gg+xlab("")+ylab("Salary")
gg


```

