---
title: "Assignment 3"
output: github_document
---

Use the attrition dataset to answer the following questions. As always, your assignment must be turned in as a .Rmd file. 

```{r}
library(tidyverse)

at<-read_csv("https://community.watsonanalytics.com/wp-content/uploads/2015/03/WA_Fn-UseC_-HR-Employee-Attrition.csv")

```


1. Create a graph that shows the distribution of monthly income.

```{r}
gg<-ggplot(data=at, aes(x=MonthlyIncome))
gg<-gg+geom_histogram()
gg
```

2. Create a graph that shows the average level of monthly income by field of education.

```{r}
gg<-ggplot(data=at,aes(x=Education,y=MonthlyIncome))
gg<-gg+geom_bar(stat="mean")
gg
```


```{r}
at_sum<-at%>%
  group_by(EducationField)%>%
  summarize(mean_income=mean(MonthlyIncome))

gg<-ggplot(data=at_sum,aes(x=fct_reorder(as.factor(EducationField),mean_income)
                                        ,y=mean_income))
gg<-gg+geom_bar(stat="identity")
gg
```


3. Create another graph that shows average level of monthly income by field of education and gender.

```{r}
at_sum<-at%>%
  group_by(EducationField,Gender)%>%
  summarize(mean_income=mean(MonthlyIncome))

gg<-ggplot(data=at_sum,aes(x=fct_reorder(as.factor(EducationField),mean_income),
                                        y=mean_income,fill=Gender))
gg<-gg+geom_bar(stat="identity",position = "dodge")
gg
```

4. Create a graph that shows average levels of monthly income by field of education, gender and job level (scale of 1-5, highest ranked employees are 5)

```{r}
at_sum<-at%>%
  group_by(EducationField,Gender,JobLevel)%>%
  summarize(mean_income=mean(MonthlyIncome))


gg<-ggplot(data=at_sum,aes(x=fct_reorder(as.factor(EducationField),mean_income),
                                        y=mean_income,fill=Gender))
gg<-gg+geom_bar(stat="identity",position = "dodge")
gg<-gg+facet_wrap(~JobLevel)
gg
```


5. Create a graph that shows average levels of montly income by field of education, gender, job level and department. 

```{r}
at_sum<-at%>%
  group_by(EducationField,Gender,JobLevel,Department)%>%
  summarize(mean_income=mean(MonthlyIncome))


gg<-ggplot(data=at_sum,aes(x=fct_reorder(as.factor(EducationField),mean_income),
                                        y=mean_income,fill=Gender))
gg<-gg+geom_bar(stat="identity",position = "dodge")
gg<-gg+facet_wrap(~Department+JobLevel,nrow=3)
gg<-gg+ylab("Average Montly Income")
gg<-gg+theme_fivethirtyeight()
gg
```

