---
title: "Assignment 5 Followup"
author: "Will Doyle"
output: github_document
---
Assignment 5
----------------

In this assignment, you'll be asked to predict two related items: first, who spends ANY money on cigarettes, according to the CEX. Second, how much do they spend? You'll need to do a few things.

1. Create a binary variable that's set to 1 if the household spent any money on cigarettes and 0 otherwise.

```{r}
library(tidyverse)
library(forcats)
library(ModelMetrics)
load("../lessons/cex.Rdata")
#binary variable for cigarette purchases
cex<-cex%>%mutate(any_cig=ifelse(is.na(cigarettes),0,1))

#Poverty recode
cex<-cex%>%mutate(pov_cym=ifelse(pov_cym=="",NA,pov_cym))
cex<-cex%>%mutate(pov_cym=fct_recode(as.factor(pov_cym),
                                     "In Poverty"="2",
                                     "Not in Poverty"="3"))
cex<-filter(cex,is.na(pov_cym)==FALSE)

#Testing/training

# Get half of the data via random sample
cex_train<-sample_frac(cex,.5)

save(cex_train,file="cex_train.Rdata")

## Testing data is the other half of the data--- the half of cex not in the 
##training dataset

cex_test<-setdiff(cex,cex_train)

save(cex_test,file="cex_test.Rdata")
##---------------------------------------
```


2. Create a model that predicts this variable. 

```{r}

mod_any_cig<-lm(any_cig~
                  inclass+
                  as.factor(educ_ref)+
                  as.factor(ref_race)+
                  as.factor(sex_ref),
                  data=cex_train)
summary(mod_any_cig)

```

3. When you're done, write some comments regarding why you think this model is a good one (write this up in paragraph form, using your finely-honed writing skills). According to this model, who spends any money at all on cigarettes?

*White men with low levels of education*

4. Create a subset of the data that only includes people who spent something (more than 0) on cigarettes. 

```{r}
cex_cigs_train<-filter(cex_train,any_cig==1)
cex_cigs_test<-filter(cex_test,any_cig==1)

```

5. Using this subset of the data, predict spending on cigarettes among the group who spent more than 0. 

```{r}
cigs_spend_mod<-lm(log(cigarettes)~
                  inclass+
                  as.factor(educ_ref)+
                  as.factor(ref_race)+
                  as.factor(sex_ref),
                  data=cex_cigs_train)
summary(cigs_spend_mod)

```

6. Make some equally well-written comments regarding which variables predict expenditures on cigarettes among people who smoke (or at least buy cigarettes). 

*White men with more money*

7. Calculate the rmse for your model agains the testing version of this data. 

```{r}
cigs_predict<-predict(cigs_spend_mod,newdata=cex_cigs_test)
exp(rmse(cigs_predict,log(cex_cigs_test$cigarettes)))
```

