---
title: "Assignment 6 followup"
output: github_document
---

```{r}
library(tidyverse)
load("../lessons/cex_test.Rdata")
```


Assignment 6
-----------------
Complete the following steps using the `cex.Rdata` or file. I want you to use the `ggplot` library and commands for all plots. 

1. Choose a dependent variable and plot that dependent variable as a function of a continuous independent variable.
2. If you need to, adjust the scale of either the dependent variable or independent variable to make more sense. 
3. Add a line of best fit to your graphic. 
4. Clean up the labeling and description of the graphic so that it could be presented to a broader audience. 
5. Now, run a regression predicting your outcome as a function of at least two independent variables. 
```{r}
cex$yvar<-cex$dine_out
cex$xvar1<-cex$inc_rank
cex$xvar2<-cex$educ_ref #must be categorical

#Create a model with two variables
mod1<-lm(yvar~
           xvar1+
           as.factor(xvar2),
         data=cex);summary(mod1)

hypo.data<-data.frame(expand.grid(xvar1=seq(min(cex$xvar1,na.rm=TRUE),
                                           max(cex$xvar1,na.rm=TRUE),
                                           length=100),
                                  xvar2=levels(cex$xvar2)))
                                          
pred_yvar<-predict(mod1,newdata=hypo.data)

plot_data<-data.frame(hypo.data,pred_yvar)

gg<-ggplot(plot_data,aes(x=xvar1,y=pred_yvar,color=xvar2))
gg<-gg+geom_smooth()



```


6. Plot your predictions. Make this plot clearly labeled and easy to read.
7. In your .Rmd file, include a few sentences describing what the graphic above shows. 