---
title: "Assignment 7 followup"
output: github_document
---

For this assignment, I want you to use the Census ACS API to download and plot data. Complete the following steps:

```{r}

library(rvest)
library(ggthemes)
library(lubridate)
library(acs)
library(noncensus)
library(leaflet)
library(tigris)
options(tigris_use_cache = TRUE) # save tigris shapefiles
library(plotly)
library(ggplot2)
library(forcats)
library(tidyverse)
```



```{r}
## Look up fips code for county
lookup_code("CA","Los Angeles") 

state_fips<-"06"
county_stub<-"037"

county_fips<-paste0(state_fips,county_stub)

# Get dataset that matches all zip codes to cities, counties and states. 
county_to_zip<-read_csv("http://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt")

#easier names to work with
names(county_to_zip)<-tolower(names(county_to_zip))

#Just zip codes in selected county
county_to_zip<-county_to_zip%>%filter(state==as.numeric(state_fips)&county==county_stub)%>%
  select(zcta5,state,county)

#list of zip codes
ziplist<-county_to_zip$zcta5

#City names
data(zip_codes)

city_zip<-zip_codes%>%filter(zip%in%ziplist)%>%select(zip,city)

#Arrange in order
city_zip<-city_zip%>%arrange(as.numeric(zip))
```


1. Download data for all of the zip codes in Los Angeles county on education levels. 


```{r}
# Get your own key and save as my_acs_key.txt
my_acs_key<-readLines("my_acs_key.txt",warn = FALSE)
acs_key<-my_acs_key

# Or just paste it here.
#acs_key<-"<your_acs_key_here>"

#List of tables: https://www.census.gov/programs-surveys/acs/technical-documentation/summary-file-documentation.html under, 1-year appendices
# b15002: education of pop over 25, by sex 
# b19001: household income over last 12 months

api.key.install(acs_key, file = "key.rda")

select_zip<-geo.make(zip.code=ziplist)

county_educ=acs.fetch(geography=select_zip,
                      endyear=2014,
                      table.number="B15002",
                      col.names="pretty",verbose=T)

acs.colnames(county_educ)

## Proprtion of individuals at college or above=
## number with college degree/
## total number
```

2. Compute the proportion of the population that has a bachelor's degree or above by zip code.
```{r}
prop_coll_above<-divide.acs(numerator=(county_educ[,15]+
                                      county_educ[,16]+
                                      county_educ[,17]+
                                      county_educ[,18]+
                                      county_educ[,32]+
                                      county_educ[,33]+
                                      county_educ[,34]+
                                      county_educ[,35]),
                            denominator=county_educ[,1]
)

```

3. Download data for all of the zip codes in LA county on family income by zip code. 
```{r}
# 19001-- family income           
county_income<-acs.fetch(geography=select_zip, 
                        endyear = 2015,
                        table.number="B19001", 
                        col.names="pretty")

acs.colnames(county_income)
```

4. Compute the proportion of the population that has family income above 75,000. 
```{r}
#Proportion above 75k-- 
prop_above_75<-divide.acs(numerator=(county_income[,13]+
                            county_income[,14]+
                            county_income[,15]+
                            county_income[,16]+
                            county_income[,17]),
                          denominator=county_income[,1]
                          )


```

5. Download data for all of the zip codes in LA county on health insurance coverage status (you'll need to look that table up here:[http://www.census.gov/programs-surveys/acs/technical-documentation/summary-file-documentation.html](http://www.census.gov/programs-surveys/acs/technical-documentation/summary-file-documentation.html).
```{r}
## Proportion with health insurance

## B27001 Health insurance
county_insurance<-acs.fetch(geography=select_zip, 
                        endyear = 2015,
                        table.number="B27001", 
                        col.names="pretty")

acs.colnames(county_insurance)
```

6. Calculate the proportion of the population in each zip code that is uninsured. 

```{r}
#Proportion with health insurance
prop_with_insurance<-divide.acs(numerator=(county_insurance[,]+
                            county_insurance[,]+
                            county_insurance[,]+
                            county_insurance[,]+
                            county_insurance[,1]),
                          denominator=county_insurance[,]
                          )
```


## Another variable of my choice: % under 25
```{r}
sex_by_age<-acs.fetch(geography=select_zip, 
                        endyear = 2015,
                        table.number="B01001", 
                        col.names="pretty")

acs.colnames(sex_by_age)
```


```{r}
                          
# Convert to tibble
county_df<-tibble(substr(geography(county_educ)[[1]],7,11),
                       as.numeric(estimate(prop_coll_above)),
                       as.numeric(estimate(prop_above_75)),
                       as.numeric(estimate(prop_with_insurance)),
                       as.numeric(estimate(prop_under_25))
                      )

# Give it easy to use names
names(county_df)<-c("zip","college_educ","income_75","insurance","youngpop")
save(county_df,file="dav.RData")

head(county_df)
```


7. Plot the proportion uninsured as a function of education, and then as a function of income. 


## Plots
```{r}
gg<-ggplot(county_df,aes(x=education,y=insurance))
gg<-gg+geom_point()+geom_smooth(method="lm")
gg
```


8. Model the proportion uninsured as a function of education, income _and one other variable of your choice_. 

```{r}
my_mod<-lm(insurance~college_educ+income_75+youngpop,data=county_df)
```


