---
title: "07-assignment"
author: "Doyle"
date: "3/30/2021"
output: html_document
---

```{r}
library(tidyverse)
library(tidycensus)
```


For this assignment, I want you to use the Census ACS API to download and plot data. Complete the following steps:

1. Download data for all of the counties in Georgia on education levels. 



2. Compute the proportion of the population in each county that has a bachelor's degree or above by county.

```{r}

my_acs_key<-readLines("~/hod_datasci_keys/my_acs_key.txt",warn = FALSE)

acs_key<-my_acs_key

# Or just paste it here.

#acs_key<-"<your_acs_key_here>"

census_api_key(acs_key)

v19 <- load_variables(2019, "acs5", cache = TRUE)


```


```{r}

var_list<-paste0("B15002_",c("001",
                            "015",
                            "016",
                            "017",
                            "018",
                            "032",
                            "033",
                            "034",
                            "035"))

educ<-get_acs(geography="county",
              state="GA",
                variables=var_list,
                output="wide",
              year = 2016
                )

names(educ)<-tolower(names(educ))

educ<-
  educ%>%
  group_by(name)%>%
  mutate(college_educ=((b15002_015ee+
                         b15002_016e+
                         b15002_017e+
                         b15002_018e+
                         b15002_032e+
                         b15002_033e+
                         b15002_034e+
                         b15002_035e)/b15002_001e)*100) %>%
    select(geoid,name,college_educ)
```

3. Download data for all of counties in Georgia for family income. 

```{r}

var_list<-paste0("B19001_",c("001",
                             "013",
                             "014",
                            "015",
                            "016",
                            "017"
                            ))

income<-get_acs("county",
                state="GA",
                variables=var_list,
                output="wide",
                year=2016
                )

names(income)<-tolower(names(income))

income<-income %>%
  mutate(
    income_75 = ((
      b19001_013e +
        b19001_014e +
        b19001_015e +
        b19001_016e +
        b19001_017e 
    ) / b19001_001e)*100
  )%>%
  select(geoid,name,income_75)


```


4. Compute the proportion of the population in each county that has family income above 75,000.

5. Download data for all of the counties in Georgia on health insurance coverages status 

```{r}

var_list<-paste0("B27001_",c("001",
                             "004",
                             "007",
                            "010",
                            "013",
                            "016",
                            "019",
                            "022",
                            "025",
                            "028",
                            "032",
                            "035",
                            "038",
                            "041",
                            "044",
                            "047",
                            "041",
                            "050",
                            "053",
                            "056"                                                                             ))

health<-get_acs("county",
                state="GA",
                variables=var_list,
                output="wide",
                year=2016
                )

names(health)<-tolower(names(health))

health<-health %>%
  mutate(
    insured = ((
b27001_004e+
  b27001_007e+
  b27001_010e+
  b27001_013e+
  b27001_016e+
  b27001_019e+
  b27001_022e+
  b27001_025e+
  b27001_028e+
  b27001_032e+
  b27001_035e+
  b27001_038e+
  b27001_041e+
  b27001_044e+
  b27001_047e+
  b27001_041e+
  b27001_050e+
  +b27001_053e+
  b27001_056e
    ) / b27001_001e)*100
  )%>%
  select(geoid,name,insured)
```


```{r}
df<-left_join(income,educ,by=c("geoid","name"))
df<-left_join(df,health,by=c("geoid","name"))
```



6. Calculate the proportion of the population in each county that is uninsured. 

```{r}

```


7. Plot the proportion uninsured as a function of education, and then as a function of income. 

```{r}
df%>%
  ggplot(aes(x=college_educ,y=insured))+
  geom_point()


df%>%
  ggplot(aes(x=income_75,y=insured))+
  geom_point()
```


8. Model the proportion uninsured as a function of education and income.


```{r}
df%>%
  lm(insured~income_75+college_educ,data=.)%>%
  summary()
```

