---
title: "Assignment 10"
author: "Will Doyle"
date: "4/20/2021"
output: html_document
---

Complete the following steps:

1. Using the counties dataset (`pd.Rdata`), create a model that predicts median household income (`median_hh_inc`).
```{r}
library(tidyverse)
library(tidymodels)
library(glmnet)
```

```{r}
load("../lessons/pd.RData")
```

```{r}
pd%>%
  ggplot(aes(x=median_hh_inc))+
  geom_density()

```


```{r}
## Define the model
lm_formula<-as.formula("median_hh_inc~
                        coll_grad_pc+
                        per_capita_inc+
                        homeown_rate")
```


```{r}

lm_rec <- recipe(lm_formula, data = pd) %>%
  step_zv(all_numeric()) %>% # drop any zero variance
  step_naomit(all_predictors())

```


## Set the kind of model to run
```{r}
lm_fit <- 
  linear_reg() %>% 
  set_engine("lm")
```

2. Provide the results of a 10-fold cross validation of your model. Describe what the results mean in a few clear sentences.


## Specify the resampling: K fold resampling, K=10
```{r}
folds <- vfold_cv(pd, v = 10)
```


## Fit Model and Cross Validate
```{r}
lm_kfold<-
  lm_fit%>%
  fit_resamples(
    lm_rec, ## Recipe: preps the data
    folds, ##resampling plan
    control=control_resamples(save_pred = TRUE)
  )
```


```{r}
lm_kfold%>%
  collect_metrics()
```


```{r}
lm_kfold%>%
  unnest(.metrics)%>%
  filter(.metric=="rmse")%>%
  ggplot(aes(x=.estimate))+
  geom_density()
```

```{r}

```




3. Provide the results of a 1000 cross fold validation with 10 percent (`p=.10`) leave out rate for the random partition. Describe what the results mean in a few clear sentences.


## Specify the resampling: bootstrap resampling
```{r}
boot <- bootstraps(pd,times = 100) ##1000 is usual minimum
```


```{r}
lm_boot<-
  lm_fit%>%
  fit_resamples(
    lm_rec, ## Recipe: preps the data
    boot, ##resampling plan
    control=control_resamples(save_pred = TRUE)
  )

```

```{r}
lm_boot%>%
  collect_metrics()
```

```{r}
lm_boot%>%
  unnest(.metrics)%>%
  filter(.metric=="rmse")%>%
  ggplot(aes(x=.estimate))+
  geom_density()
```

